### 2019-02-09 pilot_manuscript.tex
idk i reinstalled papaja??? also, small update to citations

\documentclass[,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Moral Foundations of U.S. Political News Organizations},
            pdfauthor={William E. Padfield~\& Erin M. Buchanan, Ph.D.},
            pdfkeywords={politics, morality, psycholinguistics},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Moral Foundations of U.S. Political News Organizations}
    \author{William E. Padfield\textsuperscript{1}~\& Erin M. Buchanan,
Ph.D.\textsuperscript{2}}
    \date{}

\shorttitle{MORAL NEWS}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Missouri State University\\\textsuperscript{2} }
\keywords{politics, morality, psycholinguistics}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers

\authornote{ William Padfield is a master's degree candidate
in psychology at Missouri State University. This thesis partially
fulfills the requirements for the Master of Science degree in
Psychology.

Correspondence concerning this article should be addressed to William E.
Padfield, 901 S. National Ave, Springfield, MO, 65897. E-mail:
\href{mailto:Padfield94@live.missouristate.edu}{\nolinkurl{Padfield94@live.missouristate.edu}}}

\abstract{
The media ecosystem has grown, and political opinions have diverged such
that there are competing conceptions of objective truth. Commentators
often point to political biases in news coverage as a catalyst for this
political divide. The Moral Foundations Dictionary (MFD) facilitates
identification of ideological leanings in text through frequency of the
occurrence of certain words. Through web scraping, the researchers
extracted articles from popular news sources' websites, calculated MFD
word frequencies, and identified words' respective valences. This
process attempts to uncover news outlets' positive or negative
endorsements of certain moral dimensions concomitant with a particular
ideology. In Experiment 1, the researchers gathered poltical articles
from four sources. They were unable to reveal significant differences in
moral or political endorsements, but they solidified the method to be
employed in further research. In Experiment 2, the researchers will
expand their number of sources to 10 and will analyze articles that
pertain to the 2018 confirmation hearings of U.S. Supreme Court Justice
Brett Kavanaugh. This topic was selected due to the moral disputes
associated with his nomination.

}

\begin{document}
\maketitle

In the United States, today's media landscape affords consumers a
multitude of options for obtaining political news. Since the advent of
cable news networks and the World Wide Web in the last decades of the
twentieth century, consumers have gained access to an ever-expanding
menagerie of news sources, many of which can be called up via a simple
click, touch, or swipe. Concurrent with this growth in available news
sources, concerns regarding political bias in news reporting have
entered public consciousness. For example, commentators argue that
networks including Fox News Channel and MSNBC communicate political news
from a conservative and liberal slant, respectively. These purported
biases have been a cause for concern given the potential for incomplete
or inaccurate news reporting potentially resulting from these biases.
Given the inherently moral nature of many political arguments and
positions, bias in news reporting might manifest as differing moral
appeals. Specifically, the use of differing moral language in political
articles might be an indicator of political bias in news media.

Morality and ethics have been of interest to thinkers, academics, and
philosophers since antiquity. Starting chiefly in the twentieth century,
a scientific approach to humans' understanding of morality emerged under
the domain of psychology. Theories attempting to explain the development
and application of people's moral intuitions built the foundation for
the subfield of moral psychology. As the field developed, however,
considerable debate has taken place regarding operational definitions of
\enquote{morality.} Concerns regarding operationalization remain an
issue in the field in the twenty-first century as researchers attempt to
infer moral and political leanings from text and speech.

\subsection{Moral Foundations Theory}\label{moral-foundations-theory}

As a discipline, modern moral psychology started in the late 1960s with
Lawrence Kohlberg (Haidt \& Graham, 2007). Kohlberg's research
popularized his theory of the development of moral reasoning. This
theory establishes the steps of moral reasoning through which humans
proceed as their cognitive structures assume higher levels of
sophistication and nuance (Kohlberg \& Hersh, 1977). Kohlberg borrowed
from Jean Piaget's stages of cognitive development in which children
progress from the sensorimotor through to the formal operations stage.
Similarly, Kohlberg found people typically start with a
\enquote{pre-conventional} understanding of morality during infancy in
which children understand \enquote{right} and \enquote{wrong} purely in
terms of how they interact with resultant experiences of rewards and
punishment. Typically, people progress through several steps until they
reach a \enquote{post-conventional} ethics. People who have reached the
post-conventional stage are said to be able to weigh competing
abstractions and reason their way to a conclusion that promotes justice
based upon their \enquote{self-chosen ethical principles} (Kohlberg \&
Hersh, 1977). From Kohlberg's perspective, issues of justice and
fairness comprise the foundation of morality (Haidt \& Graham, 2007).
This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psychologist
Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the
grounds that it focused solely on the moral concerns of men, and that it
ignored those of women (Haidt \& Graham, 2007). Gilligan drew attention
to purported differences in the ways men and women are taught to relate
to self and others. She offered a historic argument contending women
have traditionally filled roles related to caring and nurturing. She
pushes back against Kohlberg's assumption that moral development
replaces \enquote{rule of brute force,} as enforced by men, with the
justice-based \enquote{rule of law.} According to Gilligan, this
assumption implies women are less morally developed, owing to their
absence both in masculine displays of violence as well as in enforcement
of the law (Gilligan, 1982). Gilligan argues for the existence of a
distinct, but equal development process that women and girls must
undergo in order to develop their moral selves. Stark differences in the
ways women are traditionally taught to interact with their social world
cause them to develop ethical systems based upon their non-aggressive
relationships with others. Gilligan thus asserted morality was built
upon an alternative moral foundation: caring (Gilligan, 1982). This
debate between competing conceptions of morality did not resolve until
Gilligan and Kohlberg conceded the existence of two moral foundations:
justice and caring (Haidt \& Graham, 2007). While this new direction in
moral psychology appeared to represent a more inclusive outlook on the
construct, these novel ideas would soon be challenged on the grounds of
its apparent western-centric outlook.

Jonathan Haidt and Jesse Graham formulated Moral Foundations Theory as a
method by which to capture the entirety of humans' moral domain (Haidt
\& Graham, 2007). The researchers argued older theories of moral
psychology were focused primarily on issues of justice, fairness, and
caring - individually focused foundations of morality that align with
the beliefs of political liberals (Haidt \& Graham, 2007). In other
words, moral psychology ignored the valid moral foundations of
conservatives. Moral Foundations Theory (MFT) holds that people's moral
domain can be mapped by quantifying their endorsement of five moral
foundations: \emph{harm/care}, \emph{fairness/reciprocity},
\emph{ingroup/loyalty}, \emph{authority/respect}, and
\emph{purity/sanctity} (Haidt \& Graham, 2007).

In their brief overview of the history of moral psychology, Graham,
Haidt, and Nosek (2009) explained Shweder, Much, Mahapatra, and Park's
objections to moral psychology as it stood in the late 1980s. Their
criticism centered on the fact moral psychology concerned itself with
issues regarding justice and individuals' rights. Such a system, they
argued, did not account for moral concerns outside of the western world
(Graham et al., 2009). Individually focused concerns can be grouped
under an overarching \enquote{ethic of autonomy,} which was thought to
be one of three ethics upon which humans base moral decisions. The other
two ethics were the \enquote{ethic of community} (comprising one's duty
to their family, tribe, etc.), and the \enquote{ethic of divinity}" -
representing one's duty not to defile their God-given body and soul
(Graham et al., 2009). In the 2000s, Haidt and Graham (2007) took this
line of reasoning further in their assertion that moral psychology
favored certain political ideologies over others.

Haidt and Graham settled on these specific foundations after the
completion of a literature survey of research in anthropology and
evolutionary psychology (Graham et al., 2011). The researchers attempted
to locate virtues and morals corresponding to \enquote{evolutionary
thinking.} For instance, the researchers cited Mauss' work on reciprocal
gift-giving, which informed the establishment of the
\emph{fairness/reciprocity} foundation. Additionally, evolutionary
literature on disgust and its correlation to human behavior regarding
food and sex informed the \emph{purity/sanctity} foundation (Graham et
al., 2011). The researchers identified the five \enquote{top candidates}
for the foundations of human cultures' morality (Graham et al., 2011).

The first two foundations (\emph{harm/care} and
\emph{fairness/reciprocity}) are termed the \enquote{individualizing
foundations,} as they are centered on the concerns of individuals rather
than groups. \emph{Harm/care} represents an endorsement of compassion
and kindness, while opposing cruelty and harm.
\emph{Fairness/reciprocity} represents concerns centered on guaranteeing
individual rights as well as justice and equality among all people. The
other three foundations (\emph{ingroup/loyalty},
\emph{authority/respect}, and \emph{purity/sanctity}) are the
\enquote{binding} foundations, owing to their focus on group-related
concerns, rather than those of individuals. \emph{Ingroup/loyalty}
represents endorsements of patriotism and heroism and discourages
nonconformity and dissent. \emph{Authority/respect} represents an
endorsement of social hierarchies and traditions while denigrating
disobedience. Finally, \emph{purity/sanctity} represents concerns
regarding chastity and piety, while discouraging vices and indulgences,
including lust, avarice, and gluttony (Haidt \& Graham, 2007). Liberals
tend to endorse the individualizing foundations more than conservatives.
Conservatives, on the other hand, tend to endorse the binding
foundations more than liberals. It should be noted, however,
conservatives also tend to endorse all five foundations equally,
implying they base moral judgments on all foundations (Graham et al.,
2009).

\subsection{Moral Foundations
Dictionary}\label{moral-foundations-dictionary}

In order to capture language's role in moral and political reasoning,
Graham et al. (2009) formulated the Moral Foundations Dictionary (MFD)
in order to capture moral reasoning and justification as used in speech
and text. The MFD is composed of 259 words, with around 50 words
assigned to each of the five foundations. The researchers created a
preliminary list of words that they believed would be associated with
the five foundations. Then, using the Linguistic Inquiry and Word Count
(LIWC; Pennebaker, Booth, \& Frances, 2007) computer program, they
analyzed transcripts of liberal and conservative Christian sermons in
order to obtain frequencies of the occurrence of words from the
researchers' initial list. The researchers manually checked the results
from LIWC in order to make sure the results make sense given the
contexts and rhetorical devices used in the sermons, as word frequency
analysis ignores sentence context. The researchers offered the following
example from a Unitarian sermon as a demonstration of ambiguous
statements requiring human verification: \enquote{Don't let some
self-interested ecclesiastical or government authority tell you what to
believe, but read the Bible with your own eyes and open your heart
directly to Jesus} (Graham et al., 2009). This sentence added to the
\emph{authority/respect} total in LIWC's analysis, but it appears to
suggest that one should reject authority in this context. The
researchers eliminated this sentence from the \emph{authority/respect}
raw count on account of this discrepancy between the use of
authority-related words and the speaker's clear intentions (Graham et
al., 2009).

Similar to previous research on Moral Foundations Theory, liberal
ministers used \emph{harm}, \emph{fairness}, and \emph{ingroup} words
more often than conservative ministers. Conversely, conservative
ministers used \emph{authority} and \emph{purity} words more often than
liberal ministers. However, conservative ministers did not use
\emph{ingroup/loyalty} words more than liberals. Rather, liberal
ministers used words pertaining to \emph{ingroup/loyalty}, but in
contexts that promote rebellion and independence - causes
\emph{opposite} to positive endorsements of that foundation (Graham et
al., 2009).

To this point, most text analysis utilizing the Moral Foundations
Dictionary operationalizes endorsement of any one of the foundations as
percent occurrence of words in a given text from the foundation's
respective word list. As such, most analyses assume that zero percent
occurrence is indicative of no endorsement, while any non-zero percent
occurrence indicates endorsement of the foundation. This operational
definition may not be sufficient in describing the true nature of the
writer or speaker's endorsement of one of the sets of moral intuitions.
A quick glance at the MFD words for \emph{harm/care} reveals the
presence of words that are more closely associated with universally
accepted conceptions of \emph{harm} over \emph{care} and vice-versa
(Graham et al., 2009). For example, the word \enquote{cruel} has
relatively negative connotations compared to \enquote{benefit.} For the
\emph{harm/care} foundation, it is conceivable that use of the word
\enquote{cruel} might indicate a greater attentional focus of the idea
of \emph{harm} rather than \emph{care}.

For \emph{harm/care}, the definition of the foundation, as well as its
name, clearly distinguishes between two somewhat opposite sides of an
attentional continuum, with \emph{harm} on the negative end and
\emph{care} on the positive side. In other words, the entries in the MFD
for \emph{harm/care} have somewhat clear positive and negative valences.
The same pattern can be seen in the MFD entries for the other four
foundations. \emph{Purity/sanctity} features words that likely have a
negative valence to most observers, including \enquote{disease}" and
\enquote{trash,} along with more positive words, including
\enquote{right} and \enquote{sacred} (Graham et al., 2009). These
dichotomies, however, bring up other questions regarding the definition
and names of the other four foundations apart from \emph{harm/care}:
\emph{fairness/reciprocity}, \emph{ingroup/loyalty},
\emph{authority/respect}, and \emph{purity/sanctity}. The latter four
foundations have names that are harder to understand as a valence
continuum, as the concepts in the names are more similar, even to the
point of being virtually synonymous in the case of
\emph{fairness/reciprocity}.

When considering the issue of positive versus negative valence in MFD
words, the question of how texts are analyzed vis-a-vis the MFD remains.
How can raw percentage of MFD word occurrence capture the valence and
focus of the writer or speaker? If 2\% of a politician's speech features
positive words (i.e., \enquote{benefit} and \enquote{defend}) from the
MFD \emph{harm/care} list, how can researchers be sure the level and
nature of the speaker's \enquote{endorsement} of the foundation equals
that of another politician whose speech contained negatively connoted
MFD words from the \emph{harm/care} list? They would have equal
endorsements as far as the numbers are concerned, but the words used and
focus given are on opposite sides of the \emph{harm/care} spectrum.

This issue is compounded by the fact the Moral Foundations Questionnaire
(MFQ) and its subscales assume endorsement lies on a continuum. The
Moral Foundations Questionnaire (MFQ), which was developed subsequent to
the MFD, measures individuals' endorsements of each of the foundations
using a six-point scale (Graham et al., 2011). The questionnaire is made
up of judgment items and relevance items. Judgment items are phrased
such that the respondent signals their agreement or disagreement with
straightforward statements. An example of such a statement reads:
\enquote{It can never be right to kill another human being} (Graham et
al., 2011). Relevance items gauge the respondent's opinion regarding the
importance of foundation-related concerns. For example, the respondent
is directed to rate how important the following situation is to their
sense of morals: \enquote{whether or not someone did something
disgusting.} This example measures the relevance of the
\emph{purity/sanctity} foundations. Each foundation has a judgment and
relevance subscale, totaling 10 subscales for the MFQ (Graham et al.,
2011).

The aforementioned ambiguity of the Moral Foundations Dictionary as an
instrument becomes clearer upon closer examination of the items in the
Moral Foundations Questionnaire. One item under the
\emph{fairness/reciprocity} judgment subscale reads, \enquote{Justice is
the most important requirement for a society} (Graham et al., 2011). The
survey respondent must select a number on a scale from 1 to 6 indicating
responses spanning \enquote{strongly disagree} at 1 to \enquote{strongly
agree} at 6. While the scales in the MFQ do not represent true valence
as it pertains to individual words, it does allow for a greater degree
of specificity in terms of an individual's endorsement of a particular
moral foundation. When a respondent selects a 4 for the aforementioned
MFQ statement, they clearly are indicating they \enquote{slightly agree}
with the statement (Graham et al., 2011). This specificity is not
present in most analyses involving the MFD and percent occurrence,
unless they also take into account the valence of the words used in the
text or speech of interest.

\subsection{Valence}\label{valence}

Borrowing from Osgood's work in the 1950s, Bradley and Lang (1999)
recognized valence as one of three related dimensions comprising emotion
when developing their Affective Norms for English Words (ANEW). As
mentioned before, \enquote{valence,} the first dimension, denotes the
pleasantness of a given word. \enquote{Arousal,} the second dimension,
describes the stimulating nature of a word. Lastly, \enquote{dominance}
or \enquote{control} describes the extent to which a word makes one feel
in or out of control (Bradley \& Lang, 1999). The researchers developed
ANEW by presenting participants with a list of 100-150 words and asking
for them to rate the word on all three dimensions using the
Self-Assessment Mannikin (SAM), which allows ratings along either a
nine-point scale when using traditional paper instruments or a
twenty-point scale when using a computerized version.

Participants saw the stimulus word and responded on each scale. The
valence scale featured a smiling figure at one end (representing
pleasantness) and a frowning figure at the other end (for
unpleasantness). The arousal scale had a \enquote{wide-eyed} figure at
one end with a sleepy figure at the other, representing stimulating and
unstimulating respectively. Finally, the dominance scale featured a
large figure, indicating the highest degree of control, at one end and a
small figure, indicating a lack of control, at the other end (Bradley \&
Lang, 1999). The end result of this procedure yielded affective norms
along the three dimensions for 1,040 English words (Bradley \& Lang,
1999). ANEW represented an important first step in establishing
affective norms for large numbers of English words. However, later
researchers found the 1,040-word list to be limiting for a language
consisting of thousands of words.

Warriner, Kuperman, and Brysbaert (2013) exponentially lengthened the
list of words with affective norms to 13,915 English lemmas, the base
forms of words without inflection (i.e., \enquote{watch} rather than
\enquote{watched} and \enquote{watching}). The researchers recognized
the importance of affective norms in several areas of study, including
emotion, language processing, and memory (Warriner et al., 2013). They
argue the list of words included in ANEW is sufficient for small-scale
factorial research designs, but the list is \enquote{prohibitively
small} for larger-scale \enquote{megastudies} that are common in
psycholinguistic research today (Warriner et al., 2013).

In order to source a large number of lemmas for affective ratings, the
researchers drew from several validated sources. These include the
30,000 lemmas with age-of-acquisition (average age at which a particular
word is learned) ratings gathered by Kuperman, Stadthagen-Gonzalez, and
Brysbaert (2012) as well as the content lemmas from the SUBTLEX-US
corpus consisting of subtitles from various forms of visual media (New,
Brysbaert, Veronis, \& Pallier, 2007). This data collection resulted in
the final list of 13,915 lemmas. Lists of 346-350 words were presented
to participants recruited through the Amazon Mechanical Turk subject
pool. Participants rated the words along one of the three dimensions,
unlike the ANEW project in which participants rated each word along all
three dimensions at once. The researchers used a nine-point scale
similar to the one used by Bradley and Lang (1999) when collecting
ratings for ANEW (Warriner et al., 2013).

The researchers noted several points of interest upon observing ratings.
First, they found that valence and dominance ratings had a negative
skew, indicating more words elicited feelings of happiness and control
than their respective opposites. Also, when examining the relationship
between valence and arousal ratings, the researchers found a U-shaped
relationship. This U-shape indicates words with high degrees of
positivity and negativity elicited higher arousal (Warriner et al.,
2013). These observations along with the now-greatly expanded list of
affective norms has been applied to several lines of inquiry in
psycholinguistics.

Warriner and Kuperman (2015) utilized the new affective norms list in
order to investigate the validity of the Pollyanna hypothesis, or the
prevalence of a generally optimistic outlook in humans as reflected in
language. The researchers were able to conclude the existence of a
greater number of positive-valence English words in the list of 13,915
lemmas. Additionally, after observing token frequency in a number of
text corpora, including SUBTLEX-US, the Corpus for Contemporary American
English (COCA), the British National Corpus (BNC), Touchstone Applied
Science Associates, Inc. Corpus (TASA), and the corpus used for the
Hyperspace Analogue to Language model (HAL), the researchers found that
words with positive valence were also used more frequently (Warriner \&
Kuperman, 2015). While the researchers concede the possibility of an
acquiescence bias in ratings as a possible explanation for the observed
positivity bias, this investigation represents one application of the
Warriner et al. (2013) list in emotional studies.

In addition to applications in emotion research, the Warriner et al.
(2013) norms have been utilized in cognitive research as well. One
cognition-based study investigates the relationship between emotion and
response latencies in word recognition. Kuperman, Estes, Brysbaert, and
Warriner (2014) sought to use these new norms to fill in the knowledge
gaps regarding variance in word recognition. The researchers drew
several conclusions regarding emotion and word recognition (specifically
in naming and lexical decision tasks - two cognitive processing tasks
wherein a participant has to read aloud or judge a word for its
lexicality). First, Kuperman et al. (2014) found slower decision-making
and reading times in negative-valence words, faster times in neutral
words, and even faster times in words with positive valence. The
researchers also concluded that words causing higher arousal tend to
have slower decision times than less-arousing words. They found valence
had a stronger effect on recognition than arousal (both effects were
independent, not interactive). They found an interaction between emotion
and word frequency such that valence and arousal are more effective on
lower frequency words than high frequency words. Finally, Kuperman et
al. (2014) found a greater effect of valence and arousal on response
latency for lexical decision tasks than for naming tasks (Kuperman et
al., 2014). This research serves as further evidence that the Warriner
et al. (2013) list can be used for research inquiries both within and
without the field of psycholinguistics.

In the present studies, the researchers used the Warriner et al. (2013)
list in order to denote the valence of the words appearing in the news
articles scraped from the internet. Valence was considered as another
independent variable and its relationship with the words comprising the
Moral Foundations Dictionary were of chief interest to the researchers.
The valence was used as a means to determine whether individual words in
the MFD represented more positive aspects of their respective foundation
or if they denoted a more negative aspect of the foundation.
Specifically, valences were used to weight the MFD words by their
relative degree of positivity or negativity. Incorporating word valence
into a study involving the MFD is meant to alleviate some of the issues
regarding the aforementioned ambiguity regarding the words in the Moral
Foundations Dictionary.

\subsection{News Media and Politics}\label{news-media-and-politics}

Research into politics, language, and media has illuminated the complex
relationships between all three. Any politically-oriented discussion of
word occurrence as an implication of moral or political position assumes
that language and ideology are intrinsically linked. Deborah Cameron
(2006) points out the expressive nature of ideological beliefs and how
that expression is conveyed through language, thus implying a connection
between ideology and language. She goes on to criticize the notion that
language is either the \enquote{pre-existing raw material} used to shape
ideologies or the \enquote{post-hoc vehicle} for their propagation.
Rather, the structure of language itself is shaped by ideology and
social processes even when it is used to explain or express ideologies
(Cameron, 2006). Owing to the fact the Moral Foundations Dictionary was
developed in order to assess the moral, which includes the ideological,
orientation of discourse, its purported ability to assess parts of the
structure of language (vocabulary) for ideological lean is of chief
interest to the researchers in the present study.

The use of language both to express and further an ideological goal has
been documented in the techniques employed by candidates for political
office in the U.S., Druckman, Jacobs, and Ostermeier (2004) considered
political \enquote{issues} as communication that attempts to persuade
constituents to vote for the candidates based on their strengths in
matters of public policy. According to the researchers, \enquote{image}
priming describes techniques deployed in order to sway votes based on
favorable aspects of the candidate's behavior and personality (Druckman
et al., 2004). The researchers investigated political issue and image
priming on the part of candidates as implied by the disproportionate
attention candidates paid to particular issues over others. The
researchers found numerous examples of issue and image priming during
the 1972 re-election campaign of Richard Nixon.

They linked the Nixon administration's awareness of the issues for which
the president had public support to the issues he should emphasize (and
prime) during the campaign. Likewise the researchers found evidence that
Nixon's team was aware of negative evaluations of his warmth and
trustworthiness, and thus took steps to prime his purportedly positive
qualities, including strength and competence (Druckman et al., 2004).
The researchers also cited research from Iyengar and Kinder (1987)
suggesting the news media affected perceptions of President Jimmy
Carter's competence by emphasizing (e.g., priming) issues related to
energy, defense, and the economy. This focus implies news media may
contribute to Americans' perception of politicians based on where the
media places emphasis.

There is a potential caveat regarding the validity of Druckman et al.
(2004)'s findings: reproductions of several studies purporting to
demonstrate social priming effects have failed to replicate the original
results. Pashler, Coburn, and Harris (2012) point out the distinction
between perceptual and social (or goal) priming both in their
operational definitions as well as their replicability. Perceptual
priming often works through the inducement of a certain response from a
related prime, as in, for example, semantic priming. Social (or goal)
priming encompasses phenomena by which people exhibit complex behavioral
changes subsequent to exposure to a prime. Pashler et al. (2012) point
out well-known studies investigating social priming, including the use
of elderly-related primes to induce slower walking speeds in
participants. Studies investigating perceptual priming have been
\enquote{directly replicated in hundreds of labs} (Pashler et al.,
2012). This replication rate does not appear to be the case for social
priming, as argued by Pashler et al. (2012).

Pashler et al. (2012) noticed the unusually large effect size values
(Cohen's \emph{d}) reported by researchers studying social priming
effects. The researchers reproduced two studies from Williams and Bargh
(2008) The first study attempted to prime participants by having them
plot points on a Cartesian grid. The independent variable was priming
condition and contained three levels: short, middle, and long distance.
Those instructed to plot points further apart were hypothesized to
express a higher degree of psychological distance regarding their
family. The second study used the same priming conditions, but
hypothesized that greater distance between points would prime
participants to estimate fewer calories in unhealthy foods than those
who were primed with shorter distances between points. Pashler et al.
(2012) concluded those two studies from Williams and Bargh (2008) held
little validity while also casting doubt on the prevalence of social
priming effects themselves, based on the inability of other researchers
to replicate previously reported effects in this area.

While these concerns regarding the replication of social priming studies
are valid and deserve further investigation, Druckman et al. (2004) does
not purport to demonstrate a widespread effect of social priming on the
American electorate. In other words, this reseach makes no claim to
empirically supported priming effects. Rather, Druckman et al. (2004)
chronicle the efforts on the part of the Nixon Administration to prop up
the president's supposed strengths while downplaying his weaknesses.
These tactics were deployed through the careful use of language in order
to achieve the administration's political goals. As such, Druckman et
al. (2004)'s research on Nixon serves as an example of language's
potential utility in the propogation of desirable political opinions.
The researcher's investigation of news media's focus on specific issues
during the Carter Administration likewise provide an example of language
as a potential conduit for the transfer of politically biased
information. The idea that even 1970s news media could contain political
biases is of particular interest to the current study, which
investigates similar phenomena in contemporary news media.

Other research into news media suggests certain media outlets, at least
indirectly, may have an effect on the voting records of representatives
in Congress (Clinton \& Enamorado, 2014). Specifically, the researchers
identified a pattern of declining support for President Bill Clinton's
policies chiefly among Republicans in the House of Representatives after
the Fox News Channel began broadcasting on cable and satellite systems
in their respective districts. As Fox News was, at the time of its
launch in 1996, the only outwardly ideological national news network,
the researchers were able to track its spread across the country and
observe voting records of members of Congress both before and after Fox
News' arrival. The researchers concluded that members of Congress,
excluding those newly elected at the time of Fox News Channel's
emergence, attempted to anticipate resultant conservative-leaning shifts
among their constituents by bolstering their conservative voting record
before the next election (Clinton \& Enamorado, 2014).

Therefore, the current study sought to combine both methods related
questions and extension/replication of previous moral foundation results
found for liberal and conservative sources. First, the MFD was combined
with previous research by the current authors (see below) and weighted
by valence to create weighted percentages to better specify endorsement.
Second, these weighted percentages were examined for their differences
in across liberal and conservative news sources.

\section{Experiment 1}\label{experiment-1}

\section{Method}\label{method}

For Experiment 1, the researchers approached the study with the
intention to answer a method question. That is, this portion of the
current research was conducted in order to solidify the best method by
which to analyze political news text under the Moral Foundations Theory
framework while also alleviating some of the aforementioned valence
problem observed in the Moral Foundations Dictionary. The researchers
hypothesized the news sources genrally perceived as liberal leaning
(\emph{NPR} and \emph{The New York Times}) would contain MFD words and
valences indicating endorsements of the individualizing moral
foundations (\emph{harm/care} and \emph{fairness/reciprocity}).
Additionally, the researchers hypothesized the two sources generally
perceived to be conservative leaning (\emph{Fox News} and
\emph{Breitbart}) would feature MFD words and valences indicating equal
endorsement of all five foundations.

\subsection{Sources}\label{sources}

Political articles were collected from the websites of four notable U.S.
news sources, a process known as web scraping. The sources were
\emph{The New York Times}, \emph{National Public Radio (NPR)}, \emph{Fox
News}, and \emph{Breitbart}. They were selected for their widespread
recognition and the fact they are easily categorized (by the general
public) according to perceived political lean. In general, \emph{The New
York Times} and \emph{NPR} are perceived by many to have a liberal bias
or lean. In contrast, Fox News and Breitbart are believed to have a
conservative bias or lean (Mitchell, Matsa, Gottfried, \& Kiley, 2014).
Political articles in particular were identified and subsequently
scraped by including the specific URL directing to each source's
political content in the \emph{R} script. For example, rather than
scrape from nytimes.com, which would return undesired results
(non-political features, reviews, etc.), we instead included
nytimes.com/section/politics so that more or less exclusively political
content was obtained. All code for this manuscript can be found at
\url{https://osf.io/5kpj7/}, and the scripts are provided inline with
this manuscript written with the \emph{papaja} library (Aust \& Barth,
2017).

Identification of the sources' political URLs presented a problem for
two of the sources owing to complications with how their particular
sites were structured. While in the multi-week process of scraping
articles, we noticed word counts for \emph{NPR} and \emph{Fox News} were
not growing at a similar pace as those from \emph{The New York Times}
and \emph{Breitbart}. Upon investigation, we found another, more robust
URL for political content from NPR: their politics content
\enquote{archive.} The page structure on NPR's website was such that
only a limited selection of articles is displayed to the user at a given
time. Scraping both the archive and the normal politics page ensured we
were obtaining most (if not all) new articles as they were published. We
later ran a process in order to exclude any duplicate articles.
\emph{Fox News} presented a similar issue. We discovered \emph{Fox News}
utilized six URLs in addition to the regular politics page. These URLs
led to pages containing content pertaining the U.S. Executive Branch,
Senate, House of Representatives, Judicial Branch, foreign policy, and
elections. Once again, duplicates were subsequently eliminated from any
analyses.

\subsection{Materials}\label{materials}

Using the \emph{rvest} library in the statistical package \emph{R}, we
pulled body text for individual articles from each of the aforementioned
sources (identified using CSS language) and compiled them into a dataset
(Wickham, 2016). Using this dataset, we identified word count and
average word count per source. This process was run once daily starting
in February 2018 until March 2018. Starting in mid-March 2018, the
process was run twice daily - once in the morning and again in the
evening. Data collection was terminated once 250,000 words per source
was collected in April 2018.

\subsection{Data analysis}\label{data-analysis}

Once data collection ended, the text was scanned using the \emph{ngram}
package in \emph{R} (Schmidt, Gonzalez-Cabrera, \& Tomasello, 2017).
This package includes a word count function, which was used to remove
articles that came through as blank text, as well as to eliminate text
picked up from the Disqus commenting system used by certain websites. At
this point, duplicate articles were discarded.

The article text was processed using the \emph{tm} and \emph{ngram}
packages in \emph{R} in order to render the text in lowercase, remove
punctuation, and fix spacing issues (Feinerer \& Hornik, 2017). The
individual words were then reduced to their stems (i.e., \emph{abused}
was stemmed to \emph{abus}). The same procedure was applied to the MFD
words and the words in the Warriner et al. (2013) dataset. Using the
Warriner et al. (2013) dictionary, the words making up each of the five
foundations in the MFD were matched to their respective valence value.

Concurrent research by Jordan, Buchanan, and Padfield (2019) is
assessing the validity of both the Moral Foundations Questionnaire and
the Moral Foundations Dictionary through a multi-trait multi-method
analysis of the two instruments using multiple samples. The instruments
and foundation areas are being analyzed against one another, in order to
test reliability, as well as against the Congressional Record in order
to test predictive validity for political orientation. The researchers
were able to identify a number of potential new words that, if added to
the MFD, could comprise a dictionary with greater validity, and less
likelihood of zero percent texts, as this often occurs with the current
MFD. Those results have informed this analysis, and their updated
findings may change the underlying dictionary used in this analysis
(albeit, we do not expect any changes in the results presented below).

The source article words were compiled into a dataset where they were
matched up with their counterparts in the MFD along with their valence
and a percentage of their occurrence. Therefore, for each article, the
percentage of the number of \emph{harm/care} words occurring in the
articles were calculated, and this process was repeated for each of the
foundations. Words' percent occurence were multiplied by their
\emph{z}-scored valence. Valences were \emph{z}-scored in order to
eliminate any ambiguity regarding the direction of the valence. Positive
values indicate positive valence, and negative values indicate negative
valence. Words were categorized in accordance to their MFD affiliation,
creating a weighted sum for each moral foundation.

\section{Results}\label{results}

To analyze if news sources adhered to differences in word use based on
their target audience, we utilized a multilevel model (MLM) to analyze
the data. MLM is a regression technique that allows one to control for
the repeated measurement and nested structured of the data, which
creates correlated error (Gelman, 2006). Using the \emph{nlme} library
in \emph{R} (Pinheiro, Bates, Debroy, Sarkar, \& Team, 2017), each
foundation's weighted percentage was predicated here by the political
lean of the news source, using the individual news sources as a random
intercept to control for the structure of the data.

The multilevel model did not indicate the presence of any significant or
practical effect of political lean for any of the five moral
foundations. The strongest effect size was observed for the
\emph{authority/respect} foundation, but the effect was in the opposite
direction from what was originally hypothesized - liberal sources tended
to use more \emph{authority/respect} words than did conservative
sources. Descriptive and test statistics, \emph{p}-values and effect
sizes (Cohen's \emph{d}) can be found in Table 1. To interpret the
weighted scores, one can examine the mean and standard deviations for
each. A zero score for the mean, with a non-zero standard deviation,
would indicate a perfect balance of positive and negative words in each
category, likely representing a neutral tone when all words are
considered. Negative percentages would indicate more representation of
the negative words in the MFD area, while positive percentages indicate
an endorsement of the positive words in a MFD. Therefore, we suggest
using the sign of the mean score to determine the directionality of the
endorsement for the MFD (positive, neutral, negative), and the standard
deviation to ensure that a zero score is not zero endorsement (i.e., a
SD of zero indicates no words were used). Based on the weighted percent
values for the five foundations, the researchers observed that MFD words
seem to make up a small portion of the article text. Furthermore, the
observed percentages and means appear to indicate a generally positive
endorsement of all five foundations across both liberal and conservative
sources.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp1-table}Experiment 1 Results - Multilevel Model}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.50 & 2.21 & 0.49 & 2.21 & -0.21 & .850 & 0.01\\
Fairness/Reciprocity & 1.13 & 1.38 & 1.11 & 1.38 & -0.42 & .715 & 0.02\\
Ingroup/Loyalty & 1.28 & 1.63 & 1.34 & 1.63 & 0.30 & .789 & -0.04\\
Authority/Respect & 0.72 & 1.62 & 1.06 & 1.62 & 3.17 & .087 & -0.20\\
Purity/Sanctity & 1.11 & 1.48 & 1.27 & 1.48 & 2.37 & .141 & -0.09\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\section{Discussion}\label{discussion}

The results obtained in Experiment 1 did not confirm the hypothesis. The
researchers found little compelling evidence of an effect of partisan
lean on MFD endorsement. The strongest effect found was for the
\emph{authority/respect} foundation owing to the fact its Cohen's
\emph{d} value was greater than the other four foundations. However, the
effect was in the opposite direction of that which was hypothesized.
Specifically, the results indicated that liberal leaning sources
demonstrated higher positivity regarding that foundation than
conservatives. This result is contrary not only to the research
hypothesis for Experiment 1 but also to previous findings in Moral
Foundations Theory research. It should be noted, however, the effect
size was small and the relationship was not found to be statistically
significant.

Upon speculation, the researchers identified one possible reason for why
the results were unable to confirm the hypothesis. The selection of the
broad and amorphous topic of \enquote{political news} may have led to
the scraping of large numbers of articles with little to no
moral-centric content. Rather, many articles may have been, for example,
simple reporting on congressional procedures that would leave little
room for the use of moral language here, let alone words from the Moral
Foundations Dictionary. In short, the range of topics covered in
Experiment 1 was likely too broad. The possibility exists that a tighter
focus on one political issue or event, especially one that (on the
surface) has a stronger relationship with morality might be more
illuminating for research in moral language in news media.

Owing to the exploratory nature of Experiment 1, the researchers were
afforded the opportunity to consider changes to the method to be
utilized in Experiment 2. Generally speaking, the researchers believe
their methodology to be sound. Web scraping methods and text processing
remain viable methods for collecting large amounts of text and
subsequently rendering that text in a form suitable for data analysis.
Experiment 1 also demonstrated a method by which to address inherent
problems in the Moral Foundations Dictionary relating to valence. The
solution provided in Experiment 1 appears to provide insights into the
MFD words where none previously existed. Finally, calculating weighted
percentages and sums for each moral foundation provides an easily
interpreted summary of MFD word positivity and occurence.

While the methodology used in Experiment 1 features many strengths,
there are aspects which could be strengthened for future studies. The
researchers identified two such changes that were subsequently employed
in Experiment 2. First, the researchers elected to include more news
sources for web scraping and analysis in addition to the four used in
Experiment 1. Second, the researchers chose to focus their data
collection efforts exclusively on one event in U.S. politics: the
nomination and confirmation of Justice Brett Kavanaugh to the U.S.
Supreme Court. In Experiment 2, the researchers sought to confirm the
usefulness and validity of the method as well as test a similar
hypothesis as Experiment 1.

\section{Experiment 2}\label{experiment-2}

\subsection{Kavanaugh Supreme Court
Hearing}\label{kavanaugh-supreme-court-hearing}

In the wake of Justice Anthony Kennedy's retirement from the Supreme
Court of the United States, President Donald Trump nominated Brett
Kavanaugh as the new Associate Justice. Kavanaugh was previously on the
U.S. Court of Appeals for the District of Columbia. The Senate Judiciary
Committee began his confirmation hearing on September 4, 2018 (US
Government, 2018a). Following allegations of sexual assault by high
school classmate Dr.~Christine Blasey Ford, the committee postponed its
vote on whether or not to open the confirmation to the entire Senate.

On September 27, the committee questioned Dr.~Ford before commencing a
second round of questioning for Judge Kavanaugh (US Government, 2018b).
During the intervening weeks between hearings, two more women came
forward with two separate allegations of sexual assault on the part of
Kavanaugh. According to Nielsen reports, more than 20 million people
watched the September 27 proceedings on television (O'Connell, 2018).
This figure does not take into account viewers who watched online, nor
does it account for viewers outside the United States. On September 28,
the Senate Judiciary Committee voted to send the nomination to the
Senate floor. Senator Jeff Flake of Arizona, however, lobbied for a
week-long FBI investigation on Kavanaugh and the allegations facing him,
which the committee, and later the President, approved. The
investigation concluded with no significant findings. The Senate voted
50-48 to approve Kavanaugh's appointment on October 6, 2018 (US
Government, 2018c).

The Kavanaugh nomination, confirmation hearing, and eventual
swearing-in, as well as the news media's coverage of all three events,
feature many moral dimensions that likely differ depending on one's
morals. On one side of the debate, Kavanaugh's Supreme Court tenure
presents a prime opportunity to bring morality back into interpretation
of the Constitution. Kavanaugh's confirmation creates a conservative
stronghold among the justices on the court. Commentators have noted this
might help advance a judicial agenda that backpedals certain rights
previously upheld by the Supreme Court, including abortion and gay
marriage - social issues challenged by their opponents at least
partially on moral grounds. On the other side of the debate, the assault
allegations have energized Kavanaugh's opponents to advocate for his
rejection from the court owing to misdeeds resulting from Kavanaugh's
own alleged lack of morals. Additionally, the moral duty of the Senate
as the upper chamber in the U.S. legislature has been scrutinized in
public discourse with respect to its handling of the assault allegations
vis-a-vis Kavanaugh's confirmation.

In contrast to Experiment 1, the researchers approached Experiment 2
with the intention to confirm the method employed was valid for the
analysis of the scraped text as well as for any inferences drawn from
the analyses. For Experiment 2, the researchers hypothesized that news
sources perceived as liberal will exhibit positive endorsements of the
individualizing moral foundations (\emph{harm/care} and
\emph{fairness/reciprocity}) in their articles reporting on the
Kavanaugh confirmation hearing. News sources perceived as conservative
are hypothesized to positively endorse all five foundations equally in
their coverage of the Kavanaugh hearing. The researchers tested the
hypothesis by analyzing the content scraped from news sources' web pages
during the two weeks before and two weeks after Kavanaugh's
confirmation, owing to its prominence in the news. The content will be
analyzed for valence and moral alignment under Moral Foundations Theory.

\section{Method}\label{method-1}

\subsection{Sources}\label{sources-1}

Articles pertaining to the Brett Kavanaugh Supreme Court nomination and
confirmation were scraped from the websites of 12 U.S. news sources.
These sources were selected owing to their favorability among political
partisans according to Mitchell et al. (2014). The sources favored by
liberals were \emph{The New York Times}, \emph{National Public Radio
(NPR)}, \emph{Slate}, \emph{Huffington Post}, and \emph{Politico}
(Mitchell et al., 2014). The sources favored by conservatives included
\emph{Fox News}, \emph{Breitbart}, \emph{The Rush Limbaugh Show},
\emph{The Blaze}, and \emph{Sean Hannity}. Political articles
referencing Brett Kavanaugh's nomination process were identified and
subsequently scraped by including the URL for each source's coverage of
the nomination in the \emph{R} script. All code for this manuscript can
be found at \url{https://osf.io/5kpj7/}, and the scripts, again written
with the \emph{papaja} library in \emph{R}, are provided inline with
this manuscript (Aust \& Barth, 2017).

\subsection{Materials}\label{materials-1}

Using the \emph{rvest} library in the statistical package \emph{R}, we
pulled body text for individual articles from each of the aforementioned
10 news sources (identified using CSS language). We compiled the
articles into a dataset (Wickham, 2016). Using this dataset, we
identified word count and average word count per source. This process
was run for articles pertaining to Kavanaugh's nomination that were
published between September 13, 2018 and October 11, 2018 inclusive.
This date range was selected in reference to the widely-publicized and
viewed nomination hearing on September 27, 2018. We set the start date
at September 13 (two weeks before the hearing) and the end date at
October 11 (two weeks after the hearing) so that we could capture a
large amount of data (roughly one month) during which Kavanaugh's
nomination was at its peak saturation in news coverage.

\subsection{Data analysis}\label{data-analysis-1}

As in Experiment 1, the text was scanned with \emph{ngram}. Again, blank
articles, text from the Disqus system, and duplicate articles were
removed (Schmidt et al., 2017). The text was processed and stemmed in
order to convert to a usable form for further analysis (Feinerer \&
Hornik, 2017). Words were subsequently matched with their valences from
Warriner et al. (2013). Depending on the results of Jordan et al.
(2019)'s research, alternative forms of the Moral Foudnations Dictionary
may be imported instead of the original dictionary.

Using the \emph{tm} and \emph{ngram} packages in \emph{R}, the
researchers processed the text in order to convert it to lowercase, fix
spacing anomalies, and remove punctuation (Feinerer \& Hornik, 2017).
Each individual word was reduced to its stem (i.e., \emph{diseased} was
stemmed to \emph{diseas}). Once again, the same procedure was applied to
the MFD words and the words in the Warriner et al. (2013) dataset. Using
the Warriner et al. (2013) dictionary, the words in the MFD were
assigned their respective valence. The researchers obtained the words'
percent occurence in the text. Once again, percents were multiplied by
\emph{z}-scored valence and categorized into their proper MFD category.

\section{Experiment 2}\label{experiment-2-1}

\subsection{Results}\label{results-1}

\section{Discussion}\label{discussion-1}

\section{Conclusions}\label{conclusions}

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-Aust2017}{}
Aust, F., \& Barth, M. (2017). papaja: Create APA manuscripts with R
Markdown. Retrieved from \url{https://github.com/crsh/papaja}

\hypertarget{ref-Bradley1999}{}
Bradley, M. M., \& Lang, P. J. (1999). \emph{Affective Norms for English
Words (ANEW): Instruction Manual and Affective Ratings} (No. C-1). The
Center for Research in Psychophysiology, University of Florida.

\hypertarget{ref-Cameron2006}{}
Cameron, D. (2006). Ideology and language. \emph{Journal of Political
Ideologies}, \emph{11}(2), 141--152.
doi:\href{https://doi.org/10.1080/13569310600687916}{10.1080/13569310600687916}

\hypertarget{ref-Clinton2014}{}
Clinton, J. D., \& Enamorado, T. (2014). The National News Media's
Effect on Congress: How Fox News Affected Elites in Congress. \emph{The
Journal of Politics}, \emph{76}(4), 928--943.
doi:\href{https://doi.org/10.1017/S0022381614000425}{10.1017/S0022381614000425}

\hypertarget{ref-Druckman2004}{}
Druckman, J. N., Jacobs, L. R., \& Ostermeier, E. (2004). Candidate
Strategies to Prime Issues and Image. \emph{The Journal of Politics},
\emph{66}(4), 1180--1202.
doi:\href{https://doi.org/10.1111/j.0022-3816.2004.00295.x}{10.1111/j.0022-3816.2004.00295.x}

\hypertarget{ref-Feinerer2017}{}
Feinerer, I., \& Hornik, K. (2017). Text mining package. Retrieved from
\url{http://tm.r-forge.r-project.org/}

\hypertarget{ref-Gelman2006}{}
Gelman, A. (2006). Multilevel (hierarchical) modeling: What it can and
cannot do. \emph{Technometrics}, \emph{48}(3), 432--435.
doi:\href{https://doi.org/10.1198/004017005000000661}{10.1198/004017005000000661}

\hypertarget{ref-Gilligan1982}{}
Gilligan, C. (1982). New maps of development: New visions of maturity.
\emph{American Journal of Orthopsychiatry}, \emph{52}(2), 199--212.
doi:\href{https://doi.org/10.1111/j.1939-0025.1982.tb02682.x}{10.1111/j.1939-0025.1982.tb02682.x}

\hypertarget{ref-Graham2009}{}
Graham, J., Haidt, J., \& Nosek, B. A. (2009). Liberals and
conservatives rely on different sets of moral foundations. \emph{Journal
of Personality and Social Psychology}, \emph{96}(5), 1029--1046.
doi:\href{https://doi.org/10.1037/a0015141}{10.1037/a0015141}

\hypertarget{ref-Graham2011}{}
Graham, J., Nosek, B. A., Haidt, J., Iyer, R., Koleva, S., \& Ditto, P.
H. (2011). Mapping the moral domain. \emph{Journal of Personality and
Social Psychology}, \emph{101}(2), 366--385.
doi:\href{https://doi.org/10.1037/a0021847}{10.1037/a0021847}

\hypertarget{ref-Haidt2007}{}
Haidt, J., \& Graham, J. (2007). When Morality Opposes Justice:
Conservatives Have Moral Intuitions that Liberals may not Recognize.
\emph{Social Justice Research}, \emph{20}(1), 98--116.
doi:\href{https://doi.org/10.1007/s11211-007-0034-z}{10.1007/s11211-007-0034-z}

\hypertarget{ref-Jordan2019}{}
Jordan, K. N., Buchanan, E. M., \& Padfield, W. E. (2019). \emph{A
Validation of the Moral Foundations Questionnaire and Dictionary}.
Retrieved from \url{https://osf.io/kt9yf/}

\hypertarget{ref-Kohlberg1977}{}
Kohlberg, L., \& Hersh, R. H. (1977). Moral development: A review of the
theory. \emph{Theory into Practice}, \emph{16}(2), 53--59.
doi:\href{https://doi.org/10.1080/00405847709542675}{10.1080/00405847709542675}

\hypertarget{ref-Kuperman2014}{}
Kuperman, V., Estes, Z., Brysbaert, M., \& Warriner, A. B. (2014).
Emotion and language: Valence and arousal affect word recognition.
\emph{Journal of Experimental Psychology: General}, \emph{143}(3),
1065--1081.
doi:\href{https://doi.org/10.1037/a0035669}{10.1037/a0035669}

\hypertarget{ref-Kuperman2012}{}
Kuperman, V., Stadthagen-Gonzalez, H., \& Brysbaert, M. (2012).
Age-of-acquisition ratings for 30,000 English words. \emph{Behavior
Research Methods}, \emph{44}(4), 978--990.
doi:\href{https://doi.org/10.3758/s13428-012-0210-4}{10.3758/s13428-012-0210-4}

\hypertarget{ref-Mitchell2014}{}
Mitchell, A., Matsa, K. E., Gottfried, J., \& Kiley, J. (2014).
Political Polarization \& Media Habits \textbar{} Pew Research Center.
Retrieved from
\url{http://www.journalism.org/2014/10/21/political-polarization-media-habits/}

\hypertarget{ref-New2007}{}
New, B., Brysbaert, M., Veronis, J., \& Pallier, C. (2007). The use of
film subtitles to estimate word frequencies. \emph{Applied
Psycholinguistics}, \emph{28}(4), 661--677.
doi:\href{https://doi.org/10.1017/S014271640707035X}{10.1017/S014271640707035X}

\hypertarget{ref-OConnell2018}{}
O'Connell, M. (2018). Ford-Kavanaugh Ratings: Hearing Brings 20 Million
Viewers to Cable and Broadcast \textbar{} Hollywood Reporter. Retrieved
from
\url{https://www.hollywoodreporter.com/live-feed/ford-kavanaugh-ratings-hearing-brings-20-million-viewers-cable-broadcast-1147785}

\hypertarget{ref-Pashler2012}{}
Pashler, H., Coburn, N., \& Harris, C. R. (2012). Priming of Social
Distance? Failure to Replicate Effects on Social and Food Judgments.
\emph{PLoS ONE}, \emph{7}(8).
doi:\href{https://doi.org/10.1371/journal.pone.0042510}{10.1371/journal.pone.0042510}

\hypertarget{ref-Pennebaker2007}{}
Pennebaker, J. W., Booth, R. J., \& Frances, M. E. (2007). Liwc2007:
Linguistic inquiry and word count. Austin, TX.

\hypertarget{ref-Pinheiro2017}{}
Pinheiro, J., Bates, D., Debroy, S., Sarkar, D., \& Team, R. C. (2017).
nlme: Linear and nonlinear mixed effects models. Retrieved from
\url{https://cran.r-project.org/package=nlme}

\hypertarget{ref-Schmidt2017}{}
Schmidt, M. F., Gonzalez-Cabrera, I., \& Tomasello, M. (2017).
Children's developing metaethical judgments. \emph{Journal of
Experimental Child Psychology}, \emph{164}, 163--177.
doi:\href{https://doi.org/10.1016/j.jecp.2017.07.008}{10.1016/j.jecp.2017.07.008}

\hypertarget{ref-USGovernment2018b}{}
US Government. (2018a). Congressional Record. \emph{Congressional
Record}, \emph{164}(146), 46.
doi:\href{https://doi.org/10.1097/00017285-197507000-00018}{10.1097/00017285-197507000-00018}

\hypertarget{ref-USGovernment2018}{}
US Government. (2018b). Congressional Record. \emph{Congressional
Record}, \emph{164}(160), 93.
doi:\href{https://doi.org/10.1097/00017285-197507000-00018}{10.1097/00017285-197507000-00018}

\hypertarget{ref-USGovernment2018a}{}
US Government. (2018c). Congressional Record. \emph{Congressional
Record}, \emph{164}(167), 14.
doi:\href{https://doi.org/10.1097/00017285-197507000-00018}{10.1097/00017285-197507000-00018}

\hypertarget{ref-Warriner2015}{}
Warriner, A. B., \& Kuperman, V. (2015). Affective biases in English are
bi-dimensional. \emph{Cognition and Emotion}, \emph{29}(7), 1147--1167.
doi:\href{https://doi.org/10.1080/02699931.2014.968098}{10.1080/02699931.2014.968098}

\hypertarget{ref-Warriner2013}{}
Warriner, A. B., Kuperman, V., \& Brysbaert, M. (2013). Norms of
valence, arousal, and dominance for 13,915 English lemmas.
\emph{Behavior Research Methods}, \emph{45}(4), 1191--1207.
doi:\href{https://doi.org/10.3758/s13428-012-0314-x}{10.3758/s13428-012-0314-x}

\hypertarget{ref-Wickham2016}{}
Wickham, H. (2016). Package ` rvest '.

\hypertarget{ref-Williams2008}{}
Williams, L. E., \& Bargh, J. a. (2008). Keeping One ' s Distance.
\emph{Psychological Science}, \emph{19}(3), 302--308.
doi:\href{https://doi.org/10.1111/j.1467-9280.2008.02084.x}{10.1111/j.1467-9280.2008.02084.x}

\endgroup

\end{document}