### 2019-04-16 pilot_manuscript.tex
Knitted the thesis and saved as Padfield Thesis

\documentclass[english,,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Moral Foundations of U.S. Political News Organizations},
            pdfauthor={William E. Padfield~\& Erin M. Buchanan, Ph.D.},
            pdfkeywords={politics, morality, psycholinguistics},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Moral Foundations of U.S. Political News Organizations}
    \author{William E. Padfield\textsuperscript{1}~\& Erin M. Buchanan,
Ph.D.\textsuperscript{2}}
    \date{}

\shorttitle{MORAL NEWS}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Missouri State University\\\textsuperscript{2} Harrisburg University of Science and Technology}
\keywords{politics, morality, psycholinguistics}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers

\authornote{William Padfield is a master's degree candidate in
Psychology at Missouri State University. This thesis partially fulfills
the requirements for the Master of Science degree in Psychology.

Correspondence concerning this article should be addressed to William E.
Padfield, 901 S. National Ave, Springfield, MO, 65897. E-mail:
\href{mailto:Padfield94@live.missouristate.edu}{\nolinkurl{Padfield94@live.missouristate.edu}}}

\abstract{
The media ecosystem has grown, and political opinions have diverged such
that there are competing conceptions of objective truth. Commentators
often point to political biases in news coverage as a catalyst for this
political divide. The Moral Foundations Dictionary (MFD) facilitates
identification of ideological leanings in text through frequency of the
occurrence of certain words. Through web scraping, the researchers
extracted articles from popular news sources' websites, calculated MFD
word frequencies, and identified words' respective valences. This
process attempts to uncover news outlets' positive or negative
endorsements of certain moral dimensions concomitant with a particular
ideology. In Experiment 1, the researchers gathered poltical articles
from four sources. They were unable to reveal significant differences in
moral or political endorsements, but they solidified the method to be
employed in further research. In Experiment 2, the researchers expanded
their number of sources to 10 and analyzed articles that pertain to two
specific topics: the 2018 confirmation hearings of U.S. Supreme Court
Justice Brett Kavanaugh and the partial U.S. Goverment Shutdown of
2018-2019. Once again, no significant differences in moral or political
endorsements were found.

}

\begin{document}
\maketitle

In the United States, today's media landscape affords consumers a
multitude of options for obtaining political news. Since the advent of
cable news networks and the World Wide Web in the last decades of the
twentieth century, consumers have gained access to an ever-expanding
menagerie of news sources, many of which can be called up via a simple
click, touch, or swipe. Concurrent with this growth in available news
sources, concerns regarding political bias in news reporting have
entered public consciousness. For example, commentators argue that
networks including Fox News Channel and MSNBC communicate political news
from a conservative and liberal slant, respectively. These purported
biases have been a cause for concern given the potential for incomplete
or inaccurate news reporting potentially resulting from these biases.
Given the inherently moral nature of many political arguments and
positions, bias in news reporting might manifest as differing moral
appeals. Specifically, the use of differing moral language in political
articles might be an indicator of political bias in news media.

Morality and ethics have been of interest to thinkers, academics, and
philosophers since antiquity. Starting chiefly in the twentieth century,
a scientific approach to humans' understanding of morality emerged under
the domain of psychology. Theories attempting to explain the development
and application of people's moral intuitions built the foundation for
the subfield of moral psychology. As the field developed, however,
considerable debate has taken place regarding operational definitions of
\enquote{morality.} Concerns regarding operationalization remain an
issue in the field in the twenty-first century as researchers attempt to
infer moral and political leanings from text and speech.

\subsection{Moral Foundations Theory}\label{moral-foundations-theory}

As a discipline, modern moral psychology started in the late 1960s with
Lawrence Kohlberg (Haidt \& Graham, 2007). Kohlberg's research
popularized his theory of the development of moral reasoning. This
theory establishes the steps of moral reasoning through which humans
proceed as their cognitive structures assume higher levels of
sophistication and nuance (Kohlberg \& Hersh, 1977). Kohlberg borrowed
from Jean Piaget's stages of cognitive development in which children
progress from the sensorimotor through to the formal operations stage.
Similarly, Kohlberg found people typically start with a
\enquote{pre-conventional} understanding of morality during infancy in
which children understand \enquote{right} and \enquote{wrong} purely in
terms of how they interact with resultant experiences of rewards and
punishment. Typically, people progress through several steps until they
reach a \enquote{post-conventional} ethics. People who have reached the
post-conventional stage are said to be able to weigh competing
abstractions and reason their way to a conclusion that promotes justice
based upon their \enquote{self-chosen ethical principles} (Kohlberg \&
Hersh, 1977). From Kohlberg's perspective, issues of justice and
fairness comprise the foundation of morality (Haidt \& Graham, 2007).
This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psychologist
Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the
grounds that it focused solely on the moral concerns of men, and that it
ignored those of women (Haidt \& Graham, 2007). Gilligan drew attention
to purported differences in the ways men and women are taught to relate
to self and others. She offered a historic argument contending women
have traditionally filled roles related to caring and nurturing. She
pushes back against Kohlberg's assumption that moral development
replaces \enquote{rule of brute force,} as enforced by men, with the
justice-based \enquote{rule of law.} According to Gilligan, this
assumption implies women are less morally developed, owing to their
absence both in masculine displays of violence as well as in enforcement
of the law (Gilligan, 1982). Gilligan argues for the existence of a
distinct, but equal development process that women and girls must
undergo in order to develop their moral selves. Stark differences in the
ways women are traditionally taught to interact with their social world
cause them to develop ethical systems based upon their non-aggressive
relationships with others. Gilligan thus asserted morality was built
upon an alternative moral foundation: caring (Gilligan, 1982). This
debate between competing conceptions of morality did not resolve until
Gilligan and Kohlberg conceded the existence of two moral foundations:
justice and caring (Haidt \& Graham, 2007). While this new direction in
moral psychology appeared to represent a more inclusive outlook on the
construct, these novel ideas would soon be challenged on the grounds of
its apparent western-centric outlook.

Jonathan Haidt and Jesse Graham formulated Moral Foundations Theory as a
method by which to capture the entirety of humans' moral domain (Haidt
\& Graham, 2007). The researchers argued older theories of moral
psychology were focused primarily on issues of justice, fairness, and
caring - individually focused foundations of morality that align with
the beliefs of political liberals (Haidt \& Graham, 2007). In other
words, moral psychology ignored the valid moral foundations of
conservatives. Moral Foundations Theory (MFT) holds that people's moral
domain can be mapped by quantifying their endorsement of five moral
foundations: \emph{harm/care}, \emph{fairness/reciprocity},
\emph{ingroup/loyalty}, \emph{authority/respect}, and
\emph{purity/sanctity} (Haidt \& Graham, 2007).

In their brief overview of the history of moral psychology, Graham,
Haidt, and Nosek (2009) explained Shweder, Much, Mahapatra, and Park's
objections to moral psychology as it stood in the late 1980s. Their
criticism centered on the fact moral psychology concerned itself with
issues regarding justice and individuals' rights. Such a system, they
argued, did not account for moral concerns outside of the western world
(Graham et al., 2009). Individually focused concerns can be grouped
under an overarching \enquote{ethic of autonomy,} which was thought to
be one of three ethics upon which humans base moral decisions. The other
two ethics were the \enquote{ethic of community} (comprising one's duty
to their family, tribe, etc.), and the \enquote{ethic of divinity} -
representing one's duty not to defile their God-given body and soul
(Graham et al., 2009). In the 2000s, Haidt and Graham (2007) took this
line of reasoning further in their assertion that moral psychology
favored certain political ideologies over others.

Haidt and Graham settled on these specific foundations after the
completion of a literature survey of research in anthropology and
evolutionary psychology (Graham et al., 2011). The researchers attempted
to locate virtues and morals corresponding to \enquote{evolutionary
thinking.} For instance, the researchers cited Mauss' work on reciprocal
gift-giving, which informed the establishment of the
\emph{fairness/reciprocity} foundation. Additionally, evolutionary
literature on disgust and its correlation to human behavior regarding
food and sex informed the \emph{purity/sanctity} foundation (Graham et
al., 2011). The researchers identified the five \enquote{top candidates}
for the foundations of human cultures' morality (Graham et al., 2011).

The first two foundations (\emph{harm/care} and
\emph{fairness/reciprocity}) are termed the \enquote{individualizing
foundations,} as they are centered on the concerns of individuals rather
than groups. \emph{harm/care} represents an endorsement of compassion
and kindness, while opposing cruelty and harm.
\emph{Fairness/reciprocity} represents concerns centered on guaranteeing
individual rights as well as justice and equality among all people. The
other three foundations (\emph{ingroup/loyalty},
\emph{authority/respect}, and \emph{purity/sanctity}) are the
\enquote{binding} foundations, owing to their focus on group-related
concerns, rather than those of individuals. \emph{Ingroup/loyalty}
represents endorsements of patriotism and heroism and discourages
nonconformity and dissent. \emph{Authority/respect} represents an
endorsement of social hierarchies and traditions while denigrating
disobedience. Finally, \emph{purity/sanctity} represents concerns
regarding chastity and piety, while discouraging vices and indulgences,
including lust, avarice, and gluttony (Haidt \& Graham, 2007). Liberals
tend to endorse the individualizing foundations more than conservatives.
Conservatives, on the other hand, tend to endorse the binding
foundations more than liberals. It should be noted, however,
conservatives also tend to endorse all five foundations equally,
implying they base moral judgments on all foundations (Graham et al.,
2009).

Moral Foundations Theory has received criticism on the grounds that its
assumptions regarding moral intuitions have little empirical basis.
Suhler and Churchland (2011) list several potential weaknesses of MFT
that they argue might threaten the theory's validity. First, the authors
challenge Haidt and Graham (2007)'s claims the moral intuitions
represented by MFT are innate and modular. Suhler and Churchland (2011)
claim that advances in biological sciences (embryology and microbiology,
specifically) make it more difficult for researchers to claim any one
trait is either innate or learned through experience. Rather, behaviors
likely result from interactions between genetics and experience (Suhler
\& Churchland, 2011). According to the authors, without solid data
supporting the innateness of moral foundations, Haidt and Graham have
little from which to make such a claim. Similarly, Haidt and Graham
(2007) rely on evidence authored by evolutionary psychologists to make a
\enquote{strong modularity claim} (Suhler \& Churchland, 2011). However,
as with innateness, there is little neurobiological evidence to support
modularity.

Suhler and Churchland (2011) also criticized the content and taxonomy of
the five foundations. The authors criticize Haidt and Graham (2007)'s
omissions of additional foundations, including \emph{industry} and
\emph{modesty}, claiming these concepts are moralized in many societies
worldwide. Likewise, the authors question whether or not the foundations
are sufficiently distinct as to stand as their own foundation. For
example, Suhler and Churchland (2011) posit that \emph{ingroup/loyalty}
is merely a group-focused version of \emph{harm/care}. Work by Graham et
al. (2011) might serve to rebut this criticism, as the researchers found
their original five-factor structure seemed to best fit the data when
validating the Moral Foundations Questionnaire. Finally, Suhler and
Churchland (2011) point out that particular concepts related to a
foundation, including \enquote{anger} in \emph{fairness/reciprocity} and
\enquote{deception} in \emph{ingroup/loyalty}, could be ascribed to any
of the other foundations as well. In other words, it becomes difficult
to recognize a particular concept as indicative of any one foundation
when, in theory, they could be applied to all five (Haidt \& Graham,
2007; Suhler \& Churchland, 2011).

These criticisms of Moral Foundations Theory are valid and should be
taken into consideration when conducting research with instruments
derived from MFT. The current authors argue these criticisms are
especially valid when considered alongside questions regarding the Moral
Foundations Dictionary they state herein. However, there exists
compelling evidence regarding the validity of Moral Foundations Theory,
albeit regarding its application solely through the Moral Foundations
Questionnaire. This evidence is discussed within a brief explanation of
the Moral Foundations Questionnaire and its relationship to the Moral
Foundations Dictionary.

\subsection{Moral Foundations
Dictionary}\label{moral-foundations-dictionary}

In order to capture language's role in moral and political reasoning,
Graham et al. (2009) formulated the Moral Foundations Dictionary (MFD)
in order to capture moral reasoning and justification as used in speech
and text. The MFD is composed of 259 words, with around 50 words
assigned to each of the five foundations. The researchers created a
preliminary list of words that they believed would be associated with
the five foundations. Then, using the Linguistic Inquiry and Word Count
(LIWC; Pennebaker, Booth, \& Frances, 2007) computer program, they
analyzed transcripts of liberal and conservative Christian sermons in
order to obtain frequencies of the occurrence of words from the
researchers' initial list. The researchers manually checked the results
from LIWC in order to make sure the results make sense given the
contexts and rhetorical devices used in the sermons, as word frequency
analysis ignores sentence context. The researchers offered the following
example from a Unitarian sermon as a demonstration of ambiguous
statements requiring human verification: \enquote{Don't let some
self-interested ecclesiastical or government authority tell you what to
believe, but read the Bible with your own eyes and open your heart
directly to Jesus} (Graham et al., 2009). This sentence added to the
\emph{authority/respect} total in LIWC's analysis, but it appears to
suggest that one should reject authority in this context. The
researchers eliminated this sentence from the \emph{authority/respect}
raw count on account of this discrepancy between the use of
authority-related words and the speaker's clear intentions (Graham et
al., 2009).

Similar to previous research on Moral Foundations Theory, liberal
ministers used \emph{harm}, \emph{fairness}, and \emph{ingroup} words
more often than conservative ministers. Conversely, conservative
ministers used \emph{authority} and \emph{purity} words more often than
liberal ministers. However, conservative ministers did not use
\emph{ingroup/loyalty} words more than liberals. Rather, liberal
ministers used words pertaining to \emph{ingroup/loyalty}, but in
contexts that promote rebellion and independence - causes
\emph{opposite} to positive endorsements of that foundation (Graham et
al., 2009).

To this point, most text analysis utilizing the Moral Foundations
Dictionary operationalizes endorsement of any one of the foundations as
percent occurrence of words in a given text from the foundation's
respective word list. As such, most analyses assume that zero percent
occurrence is indicative of no endorsement, while any non-zero percent
occurrence indicates endorsement of the foundation. This operational
definition may not be sufficient in describing the true nature of the
writer or speaker's endorsement of one of the sets of moral intuitions.
A quick glance at the MFD words for \emph{harm/care} reveals the
presence of words that are more closely associated with universally
accepted conceptions of \emph{harm} over \emph{care} and vice-versa
(Graham et al., 2009). For example, the word \enquote{cruel} has
relatively negative connotations compared to \enquote{benefit.} For the
\emph{harm/care} foundation, it is conceivable that use of the word
\enquote{cruel} might indicate a greater attentional focus of the idea
of \emph{harm} rather than \emph{care}.

For \emph{harm/care}, the definition of the foundation, as well as its
name, clearly distinguishes between two somewhat opposite sides of an
attentional continuum, with \emph{harm} on the negative end and
\emph{care} on the positive side. In other words, the entries in the MFD
for \emph{harm/care} have somewhat clear positive and negative valences.
The same pattern can be seen in the MFD entries for the other four
foundations. \emph{Purity/sanctity} features words that likely have a
negative valence to most observers, including \enquote{disease} and
\enquote{trash,} along with more positive words, including
\enquote{right} and \enquote{sacred} (Graham et al., 2009). These
dichotomies, however, bring up other questions regarding the definition
and names of the other four foundations apart from \emph{harm/care}:
\emph{fairness/reciprocity}, \emph{ingroup/loyalty},
\emph{authority/respect}, and \emph{purity/sanctity}. The latter four
foundations have names that are harder to understand as a valence
continuum, as the concepts in the names are more similar, even to the
point of being virtually synonymous in the case of
\emph{fairness/reciprocity}.

When considering the issue of positive versus negative valence in MFD
words, the question of how texts are analyzed vis-a-vis the MFD remains.
How can raw percentage of MFD word occurrence capture the valence and
focus of the writer or speaker? If 2\% of a politician's speech features
positive words (i.e., \enquote{benefit} and \enquote{defend}) from the
MFD \emph{harm/care} list, how can researchers be sure the level and
nature of the speaker's \enquote{endorsement} of the foundation equals
that of another politician whose speech contained negatively connoted
MFD words from the \emph{harm/care} list? They would have equal
endorsements as far as the numbers are concerned, but the words used and
focus given are on opposite sides of the \emph{harm/care} spectrum.

This issue is compounded by the fact the Moral Foundations Questionnaire
(MFQ) and its subscales assume endorsement lies on a continuum. The MFQ,
which was developed subsequent to the MFD, measures individuals'
endorsements of each of the foundations using a six-point scale (Graham
et al., 2011). The questionnaire is made up of judgment items and
relevance items. Judgment items are phrased such that the respondent
signals their agreement or disagreement with straightforward statements.
An example of such a statement reads: \enquote{It can never be right to
kill another human being} (Graham et al., 2011). Relevance items gauge
the respondent's opinion regarding the importance of foundation-related
concerns. For example, the respondent is directed to rate how important
the following situation is to their sense of morals: \enquote{whether or
not someone did something disgusting.} This example measures the
relevance of the \emph{purity/sanctity} foundations. Each foundation has
a judgment and relevance subscale, totaling 10 subscales for the MFQ
(Graham et al., 2011).

The Moral Foundations Questionnaire has been validated by multiple
researchers. Likewise, its five-factor structure has been demonstrated
to fit data in multiple countries (Davies, Sibley, \& Liu, 2014; Graham
et al., 2011). In their article introducing the 30-item MFQ, Graham et
al. (2011) conducted an exploratory factor analysis (EFA) and found that
a five-factor model fit the data better than a one, two, or three-factor
model. Davies et al. (2014) conducted a confirmatory factor analysis
(CFA) of the MFQ with a sample from New Zealand and likewise found the
five-factor model provided the best fit. While Davies et al. (2014)
concede the US and New Zealand share many similarities as Western
nations, which could raise questions regarding the validity of the MFQ
in non-Western nations. However, there are striking differences between
the two countries, including the lack of a two-party political system in
New Zealand, that provide grounds for claiming the MFQ generalizes
beyond the United States. Furthermore, Graham et al. (2011) claim to
find a \enquote{reasonable} degree of generalizability for the MFQ
across participants from many different regions in the world. These two
bodies of work also provide the best available evidence that the five
moral foundations are sufficiently distinct from one another, though
broader criticisms of MFT raised by Suhler and Churchland (2011) should
still be taken into account in studies involving the theory.

The aforementioned ambiguity of the Moral Foundations Dictionary as an
instrument becomes clearer upon closer examination of the items in the
Moral Foundations Questionnaire. One item under the
\emph{fairness/reciprocity} judgment subscale reads, \enquote{Justice is
the most important requirement for a society} (Graham et al., 2011). The
survey respondent must select a number on a scale from 1 to 6 indicating
responses spanning \enquote{strongly disagree} at 1 to \enquote{strongly
agree} at 6. While the scales in the MFQ do not represent true valence
as it pertains to individual words, it does allow for a greater degree
of specificity in terms of an individual's endorsement of a particular
moral foundation. When a respondent selects a 4 for the aforementioned
MFQ statement, they clearly are indicating they \enquote{slightly agree}
with the statement (Graham et al., 2011). This specificity is not
present in most analyses involving the MFD and percent occurrence,
unless they also take into account the valence of the words used in the
text or speech of interest.

\subsection{Valence}\label{valence}

Borrowing from Osgood's work in the 1950s, Bradley and Lang (1999)
recognized valence as one of three related dimensions comprising emotion
when developing their Affective Norms for English Words (ANEW). As
mentioned before, \enquote{valence,} the first dimension, denotes the
pleasantness of a given word. \enquote{Arousal,} the second dimension,
describes the stimulating nature of a word. Lastly, \enquote{dominance}
or \enquote{control} describes the extent to which a word makes one feel
in or out of control (Bradley \& Lang, 1999). The researchers developed
ANEW by presenting participants with a list of 100-150 words and asking
for them to rate the word on all three dimensions using the
Self-Assessment Mannikin (SAM), which allows ratings along either a
nine-point scale when using traditional paper instruments or a
twenty-point scale when using a computerized version.

Participants saw the stimulus word and responded on each scale. The
valence scale featured a smiling figure at one end (representing
pleasantness) and a frowning figure at the other end (for
unpleasantness). The arousal scale had a \enquote{wide-eyed} figure at
one end with a sleepy figure at the other, representing stimulating and
unstimulating respectively. Finally, the dominance scale featured a
large figure, indicating the highest degree of control, at one end and a
small figure, indicating a lack of control, at the other end (Bradley \&
Lang, 1999). The end result of this procedure yielded affective norms
along the three dimensions for 1,040 English words (Bradley \& Lang,
1999). ANEW represented an important first step in establishing
affective norms for large numbers of English words. However, later
researchers found the 1,040-word list to be limiting for a language
consisting of thousands of words.

Warriner, Kuperman, and Brysbaert (2013) exponentially lengthened the
list of words with affective norms to 13,915 English lemmas, the base
forms of words without inflection (i.e., \enquote{watch} rather than
\enquote{watched} and \enquote{watching}). The researchers recognized
the importance of affective norms in several areas of study, including
emotion, language processing, and memory (Warriner et al., 2013). They
argue the list of words included in ANEW is sufficient for small-scale
factorial research designs, but the list is \enquote{prohibitively
small} for larger-scale \enquote{megastudies} that are common in
psycholinguistic research today (Warriner et al., 2013).

In order to source a large number of lemmas for affective ratings, the
researchers drew from several validated sources. These include the
30,000 lemmas with age-of-acquisition (average age at which a particular
word is learned) ratings gathered by Kuperman, Stadthagen-Gonzalez, and
Brysbaert (2012) as well as the content lemmas from the SUBTLEX-US
corpus consisting of subtitles from various forms of visual media (New,
Brysbaert, Veronis, \& Pallier, 2007). This data collection resulted in
the final list of 13,915 lemmas. Lists of 346-350 words were presented
to participants recruited through the Amazon Mechanical Turk subject
pool. Participants rated the words along one of the three dimensions,
unlike the ANEW project in which participants rated each word along all
three dimensions at once. The researchers used a nine-point scale
similar to the one used by Bradley and Lang (1999) when collecting
ratings for ANEW (Warriner et al., 2013).

The researchers noted several points of interest upon observing ratings.
First, they found that valence and dominance ratings had a negative
skew, indicating more words elicited feelings of happiness and control
than their respective opposites. Also, when examining the relationship
between valence and arousal ratings, the researchers found a U-shaped
relationship. This U-shape indicates words with high degrees of
positivity and negativity elicited higher arousal (Warriner et al.,
2013). These observations along with the now-greatly expanded list of
affective norms has been applied to several lines of inquiry in
psycholinguistics.

Warriner and Kuperman (2015) utilized the new affective norms list in
order to investigate the validity of the Pollyanna hypothesis, or the
prevalence of a generally optimistic outlook in humans as reflected in
language. The researchers were able to conclude the existence of a
greater number of positive-valence English words in the list of 13,915
lemmas. Additionally, after observing token frequency in a number of
text corpora, including SUBTLEX-US, the Corpus for Contemporary American
English (COCA), the British National Corpus (BNC), Touchstone Applied
Science Associates, Inc. Corpus (TASA), and the corpus used for the
Hyperspace Analogue to Language model (HAL), the researchers found that
words with positive valence were also used more frequently (Warriner \&
Kuperman, 2015). While the researchers concede the possibility of an
acquiescence bias in ratings as a possible explanation for the observed
positivity bias, this investigation represents one application of the
Warriner et al. (2013) list in emotional studies.

In addition to applications in emotion research, the Warriner et al.
(2013) norms have been utilized in cognitive research as well. One
cognition-based study investigates the relationship between emotion and
response latencies in word recognition. Kuperman, Estes, Brysbaert, and
Warriner (2014) sought to use these new norms to fill in the knowledge
gaps regarding variance in word recognition. The researchers drew
several conclusions regarding emotion and word recognition (specifically
in naming and lexical decision tasks - two cognitive processing tasks
wherein a participant has to read aloud or judge a word for its
lexicality). First, Kuperman et al. (2014) found slower decision-making
and reading times in negative-valence words, faster times in neutral
words, and even faster times in words with positive valence. The
researchers also concluded that words causing higher arousal tend to
have slower decision times than less-arousing words. They found valence
had a stronger effect on recognition than arousal (both effects were
independent, not interactive). They found an interaction between emotion
and word frequency such that valence and arousal are more effective on
lower frequency words than high frequency words. Finally, Kuperman et
al. (2014) found a greater effect of valence and arousal on response
latency for lexical decision tasks than for naming tasks (Kuperman et
al., 2014). This research serves as further evidence that the Warriner
et al. (2013) list can be used for research inquiries both within and
without the field of psycholinguistics.

In the present studies, the researchers used the Warriner et al. (2013)
list in order to denote the valence of the words appearing in the news
articles scraped from the internet. Valence was considered as another
independent variable and its relationship with the words comprising the
Moral Foundations Dictionary were of chief interest to the researchers.
The valence was used as a means to determine whether individual words in
the MFD represented more positive aspects of their respective foundation
or if they denoted a more negative aspect of the foundation.
Specifically, valences were used to weight the MFD words by their
relative degree of positivity or negativity. Incorporating word valence
into a study involving the MFD is meant to alleviate some of the issues
regarding the aforementioned ambiguity regarding the words in the Moral
Foundations Dictionary.

\subsection{News Media and Politics}\label{news-media-and-politics}

Research into politics, language, and media has illuminated the complex
relationships between all three. Any politically-oriented discussion of
word occurrence as an implication of moral or political position assumes
that language and ideology are intrinsically linked. Deborah Cameron
(2006) points out the expressive nature of ideological beliefs and how
that expression is conveyed through language, thus implying a connection
between ideology and language. She goes on to criticize the notion that
language is either the \enquote{pre-existing raw material} used to shape
ideologies or the \enquote{post-hoc vehicle} for their propagation.
Rather, the structure of language itself is shaped by ideology and
social processes even when it is used to explain or express ideologies
(Cameron, 2006). Owing to the fact the Moral Foundations Dictionary was
developed in order to assess the moral, which includes the ideological,
orientation of discourse, its purported ability to assess parts of the
structure of language (vocabulary) for ideological lean is of chief
interest to the researchers in the present study.

The use of language both to express and further an ideological goal has
been documented in the techniques employed by candidates for political
office in the U.S., Druckman, Jacobs, and Ostermeier (2004) considered
political \enquote{issues} as communication that attempts to persuade
constituents to vote for the candidates based on their strengths in
matters of public policy. According to the researchers, \enquote{image}
priming describes techniques deployed in order to sway votes based on
favorable aspects of the candidate's behavior and personality (Druckman
et al., 2004). The researchers investigated political issue and image
priming on the part of candidates as implied by the disproportionate
attention candidates paid to particular issues over others. The
researchers found numerous examples of issue and image priming during
the 1972 re-election campaign of Richard Nixon.

They linked the Nixon administration's awareness of the issues for which
the president had public support to the issues he should emphasize (and
prime) during the campaign. Likewise the researchers found evidence that
Nixon's team was aware of negative evaluations of his warmth and
trustworthiness, and thus took steps to prime his purportedly positive
qualities, including strength and competence (Druckman et al., 2004).
The researchers also cited research from Iyengar and Kinder (1987)
suggesting the news media affected perceptions of President Jimmy
Carter's competence by emphasizing (e.g., priming) issues related to
energy, defense, and the economy. This focus implies news media may
contribute to Americans' perception of politicians based on where the
media places emphasis.

There is a potential caveat regarding the validity of Druckman et al.
(2004)'s findings: reproductions of several studies purporting to
demonstrate social priming effects have failed to replicate the original
results. Pashler, Coburn, and Harris (2012) point out the distinction
between perceptual and social (or goal) priming both in their
operational definitions as well as their replicability. Perceptual
priming often works through the inducement of a certain response from a
related prime, as in, for example, semantic priming. Social (or goal)
priming encompasses phenomena by which people exhibit complex behavioral
changes subsequent to exposure to a prime. Pashler et al. (2012) point
out well-known studies investigating social priming, including the use
of elderly-related primes to induce slower walking speeds in
participants. Studies investigating perceptual priming have been
\enquote{directly replicated in hundreds of labs} (Pashler et al.,
2012). This replication rate does not appear to be the case for social
priming, as argued by Pashler et al. (2012).

Pashler et al. (2012) noticed the unusually large effect size values
(Cohen's \emph{d}) reported by researchers studying social priming
effects. The researchers reproduced two studies from Williams and Bargh
(2008) The first study attempted to prime participants by having them
plot points on a Cartesian grid. The independent variable was priming
condition and contained three levels: short, middle, and long distance.
Those instructed to plot points further apart were hypothesized to
express a higher degree of psychological distance regarding their
family. The second study used the same priming conditions, but
hypothesized that greater distance between points would prime
participants to estimate fewer calories in unhealthy foods than those
who were primed with shorter distances between points. Pashler et al.
(2012) concluded those two studies from Williams and Bargh (2008) held
little validity while also casting doubt on the prevalence of social
priming effects themselves, based on the inability of other researchers
to replicate previously reported effects in this area.

While these concerns regarding the replication of social priming studies
are valid and deserve further investigation, Druckman et al. (2004) does
not purport to demonstrate a widespread effect of social priming on the
American electorate. In other words, this research makes no claim to
empirically supported priming effects. Rather, Druckman et al. (2004)
chronicle the efforts on the part of the Nixon Administration to prop up
the president's supposed strengths while downplaying his weaknesses.
These tactics were deployed through the careful use of language in order
to achieve the administration's political goals. As such, Druckman et
al. (2004)'s research on Nixon serves as an example of language's
potential utility in the propagation of desirable political opinions.
The researcher's investigation of news media's focus on specific issues
during the Carter Administration likewise provide an example of language
as a potential conduit for the transfer of politically biased
information. The idea that even 1970s news media could contain political
biases is of particular interest to the current study, which
investigates similar phenomena in contemporary news media.

Other research into news media suggests certain media outlets, at least
indirectly, may have an effect on the voting records of representatives
in Congress (Clinton \& Enamorado, 2014). Specifically, the researchers
identified a pattern of declining support for President Bill Clinton's
policies chiefly among Republicans in the House of Representatives after
the Fox News Channel began broadcasting on cable and satellite systems
in their respective districts. As Fox News was, at the time of its
launch in 1996, the only outwardly ideological national news network,
the researchers were able to track its spread across the country and
observe voting records of members of Congress both before and after Fox
News' arrival. The researchers concluded that members of Congress,
excluding those newly elected at the time of Fox News Channel's
emergence, attempted to anticipate resultant conservative-leaning shifts
among their constituents by bolstering their conservative voting record
before the next election (Clinton \& Enamorado, 2014).

Therefore, the current study sought to combine both methods related
questions and extension/replication of previous moral foundation results
found for liberal and conservative sources. First, the MFD was combined
with previous research by the current authors (see below) and weighted
by valence to create weighted percentages to better specify endorsement.
Second, these weighted percentages were examined for their differences
in across liberal and conservative news sources.

\section{Experiment 1}\label{experiment-1}

\section{Method}\label{method}

For Experiment 1, the researchers approached the study with the
intention to answer a method question. That is, this portion of the
current research was conducted in order to solidify the best method by
which to analyze political news text under the Moral Foundations Theory
framework while also alleviating some of the aforementioned valence
problem observed in the Moral Foundations Dictionary. The researchers
hypothesized the news sources generally perceived as liberal leaning
(\emph{NPR} and \emph{The New York Times}) would contain MFD words and
valences indicating endorsements of the individualizing moral
foundations (\emph{harm/care} and \emph{fairness/reciprocity}).
Additionally, the researchers hypothesized the two sources generally
perceived to be conservative leaning (\emph{Fox News} and
\emph{Breitbart}) would feature MFD words and valences indicating equal
endorsement of all five foundations. Owing to the lack of a need for
human participants, the researchers did not petition Missouri State
University's Institutional Review Board, as no such approval was needed
to conduct this study.

\subsection{Sources}\label{sources}

Political articles were collected from the websites of four notable U.S.
news sources, a process known as web scraping. The sources were
\emph{The New York Times}, \emph{National Public Radio (NPR)}, \emph{Fox
News}, and \emph{Breitbart}. They were selected for their widespread
recognition and the fact political partisans have strong preferences for
some sources over others. The researchers determined the political lean
of each source by referencing Mitchell, Matsa, Gottfried, and Kiley
(2014)'s article demonstrating the self-reported ideological consistency
represented by the consumers of several news sources. In general,
\emph{The New York Times} and \emph{NPR} are preferred by consumers
reporting a liberal bias or lean. In contrast, \emph{Fox News} and
\emph{Breitbart} are believed to have a conservative bias or lean.
Mitchell et al. (2014)'s article presented political ideology as a scale
ranging from \enquote{consistently liberal} to \enquote{consistently
conservative.} In between these extremes lie more moderate positions,
including \enquote{mostly liberal,} \enquote{mixed,} and \enquote{mostly
conservative.} Owing to the lower number of sources analyzed herein, the
researchers elected to categorize the sources as either
\enquote{liberal} and \enquote{conservative} in order to form a basis
for comparison.

Political articles in particular were identified and subsequently
scraped by including the specific URL directing to each source's
political content in the \emph{R} script. For example, rather than
scrape from nytimes.com, which would return undesired results
(non-political features, reviews, etc.), we instead included
nytimes.com/section/politics so that more or less exclusively political
content was obtained. All code for this manuscript can be found at
\url{https://osf.io/5kpj7/}, and the scripts are provided inline with
this manuscript written with the \emph{papaja} library (Aust \& Barth,
2017).

Identification of the sources' political URLs presented a problem for
two of the sources owing to complications with how their particular
sites were structured. While in the multi-week process of scraping
articles, we noticed word counts for \emph{NPR} and \emph{Fox News} were
not growing at a similar pace as those from \emph{The New York Times}
and \emph{Breitbart}. Upon investigation, we found another, more robust
URL for political content from NPR: their politics content
\enquote{archive.} The page structure on NPR's website was such that
only a limited selection of articles is displayed to the user at a given
time. Scraping both the archive and the normal politics page ensured we
were obtaining most (if not all) new articles as they were published. We
later ran a process in order to exclude any duplicate articles.
\emph{Fox News} presented a similar issue. We discovered \emph{Fox News}
utilized six URLs in addition to the regular politics page. These URLs
led to pages containing content pertaining the U.S. Executive Branch,
Senate, House of Representatives, Judicial Branch, foreign policy, and
elections. Once again, duplicates were subsequently eliminated from any
analyses.

\subsection{Materials}\label{materials}

Using the \emph{rvest} library in the statistical package \emph{R}, we
pulled body text for individual articles from each of the aforementioned
sources (identified using CSS language) and compiled them into a dataset
(Wickham, 2016). Using this dataset, we identified word count and
average word count per source. This process was run once daily starting
in February 2018 until March 2018. Starting in mid-March 2018, the
process was run twice daily - once in the morning and again in the
evening. Data collection was terminated once 250,000 words per source
was collected in April 2018.

\subsection{Data analysis}\label{data-analysis}

Once data collection ended, the text was scanned using the \emph{ngram}
package in \emph{R} (Schmidt, Gonzalez-Cabrera, \& Tomasello, 2017).
This package includes a word count function, which was used to remove
articles that came through as blank text, as well as to eliminate text
picked up from the Disqus commenting system used by certain websites. At
this point, duplicate articles were discarded.

The article text was processed using the \emph{tm} and \emph{ngram}
packages in \emph{R} in order to render the text in lowercase, remove
punctuation, and fix spacing issues (Feinerer \& Hornik, 2017). The
individual words were then reduced to their stems (i.e., \emph{abused}
was stemmed to \emph{abus}). The same procedure was applied to the MFD
words and the words in the Warriner et al. (2013) dataset. Using the
Warriner et al. (2013) dictionary, the words making up each of the five
foundations in the MFD were matched to their respective valence value.

Concurrent research by Jordan, Buchanan, and Padfield (2019) is
assessing the validity of both the Moral Foundations Questionnaire and
the Moral Foundations Dictionary through a multi-trait multi-method
analysis of the two instruments using multiple samples. The instruments
and foundation areas are being analyzed against one another, in order to
test reliability, as well as against the Congressional Record in order
to test predictive validity for political orientation. The researchers
were able to identify a number of potential new words that, if added to
the MFD, could comprise a dictionary with greater validity, and less
likelihood of zero percent texts, as this often occurs with the current
MFD. Those results have informed this analysis, and their updated
findings may change the underlying dictionary used in this analysis
(albeit, we do not expect any changes in the results presented below).

The source article words were compiled into a dataset where they were
matched up with their counterparts in the MFD along with their valence
and a percentage of their occurrence. Therefore, for each article, the
percentage of the number of \emph{harm/care} words occurring in the
articles were calculated, and this process was repeated for each of the
foundations. Words' percent occurrence were multiplied by their
\emph{z}-scored valence. Valences were \emph{z}-scored in order to
eliminate any ambiguity regarding the direction of the valence. Positive
values indicate positive valence, and negative values indicate negative
valence. Words were categorized in accordance to their MFD affiliation,
creating a weighted sum for each moral foundation.

\section{Results}\label{results}

\subsection{Descriptive Statistics}\label{descriptive-statistics}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp1-source-descriptives}Experiment 1 - Descriptive Statistics by Source}
\begin{tabular}{lcccccclccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
NPR & 0.28 & 0.23 & 695 & 302977 & 435.94 & 642.63 & 191.96 & 192.29 & 13.80 & 3.93\\
New York Times & 0.30 & 0.13 & 406 & 452579 & 1,114.73 & 511.86 & 454.27 & 154.58 & 16.32 & 3.29\\
Breitbart & 0.29 & 0.18 & 1437 & 722022 & 502.45 & 347.90 & 243.36 & 120.76 & 18.42 & 8.01\\
Fox News & 0.29 & 0.17 & 503 & 296779 & 590.02 & 528.60 & 283.57 & 189.00 & 16.92 & 7.16\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Readability statistics were calculated using the Flesch-Kincaid Grade Level readability formula. V = Valence, T = Tokens or total words, Ty = Types or unique words, FK = Flesch-Kincaid}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

The researchers calculated descriptive statistics for each news source
in order to understand any and all fundamental linguistic differences in
the sources' use of English. Statistics calculated included average
\emph{z}-scored valence of the unique words per article, number of
articles per source, total number of words per source, average number of
tokens (words) per article in each source, average number of types
(unique words) per article in each source, and mean readability level
per source. Readability statistics were calculated using the
Flesch-Kincaid Grade Level Readability formula (Kincaid, Fishburne,
Robert P., Richard L., \& Brad S., 1975). Readability is calculated
using a formula where the total number of syllables, words, and
sentences in a given passage are determinants of its difficulty. The
obtained value is intended to match up with the grade level at which one
should be able to comfortably read the passage (Kincaid et al., 1975).
For example, a text with a readability score of 11 should be easily read
by a high school junior.

As seen in Table \ref{tab:exp1-source-descriptives}, the sources are
similar in some aspects yet different in others. Valence appears to be
slightly positive across all sources. The large standard deviations seem
to indicate little to no presence of a difference in valence. \emph{The
New York Times} published the greatest number of articles as well as
total words. \emph{Breitbart} featured the lowest number of articles,
and \emph{NPR} the lowest number of total words from all articles. Per
individual article, however, \emph{Breitbart} appears to feature the
highest average number of words as well as unique words. Once again the
standard deviations call into question any apparent differences between
sources. Finally, \emph{Fox News} appears to be the most readable source
on average, while \emph{The New York Times} is the least readable. This
result might be attributable to the greater number of tokens in the
average \emph{New York Times} article compared to \emph{Fox News}. As
before, the standard deviations for readability hamper one's ability to
draw solid conclusions regarding differences in the sources' use of
language.

\subsection{Inferential Statistics}\label{inferential-statistics}

To analyze if news sources adhered to differences in word use based on
their target audience, we utilized a multilevel model (MLM) to analyze
the data. MLM is a regression technique that allows one to control for
the repeated measurement and nested structured of the data, which
creates correlated error (Gelman, 2006). Using the \emph{nlme} library
in \emph{R} (Pinheiro, Bates, Debroy, Sarkar, \& Team, 2017), each
foundation's weighted percentage was predicated here by the political
lean of the news source, using the individual news sources as a random
intercept to control for the structure of the data.

The multilevel model did not indicate the presence of any significant or
practical effect of political lean for any of the five moral
foundations. The strongest effect size was observed for the
\emph{authority/respect} foundation, but the effect was in the opposite
direction from what was originally hypothesized - liberal sources tended
to use more \emph{authority/respect} words than did conservative
sources. Descriptive and test statistics, \emph{p}-values and effect
sizes (Cohen's \emph{d}) can be found in Table \ref{tab:exp1-table}. To
interpret the weighted scores, one can examine the mean and standard
deviations for each. A zero score for the mean, with a non-zero standard
deviation, would indicate a perfect balance of positive and negative
words in each category, likely representing a neutral tone when all
words are considered. Negative percentages would indicate more
representation of the negative words in the MFD area, while positive
percentages indicate an endorsement of the positive words in a MFD.
Therefore, we suggest using the sign of the mean score to determine the
directionality of the endorsement for the MFD (positive, neutral,
negative), and the standard deviation to ensure that a zero score is not
zero endorsement (i.e., a SD of zero indicates no words were used).
Based on the weighted percent values for the five foundations, the
researchers observed that MFD words seem to make up a small portion of
the article text. Furthermore, the observed percentages and means appear
to indicate a generally positive endorsement of all five foundations
across both liberal and conservative sources.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp1-table}}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.50 & 2.21 & 0.49 & 2.21 & -0.21 & .850 & 0.01\\
Fairness/Reciprocity & 1.13 & 1.38 & 1.11 & 1.38 & -0.42 & .715 & 0.02\\
Ingroup/Loyalty & 1.28 & 1.63 & 1.34 & 1.63 & 0.30 & .789 & -0.04\\
Authority/Respect & 0.72 & 1.62 & 1.06 & 1.62 & 3.17 & .087 & -0.20\\
Purity/Sanctity & 1.11 & 1.48 & 1.27 & 1.48 & 2.37 & .141 & -0.09\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\section{Discussion}\label{discussion}

The results obtained in Experiment 1 did not confirm the hypothesis. The
researchers found little compelling evidence of an effect of partisan
lean on MFD endorsement. The strongest effect found was for the
\emph{authority/respect} foundation owing to the fact its Cohen's
\emph{d} value was greater than the other four foundations. However, the
effect was in the opposite direction of that which was hypothesized.
Specifically, the results indicated that liberal leaning sources
demonstrated higher positivity regarding that foundation than
conservatives. This result is contrary not only to the research
hypothesis for Experiment 1 but also to previous findings in Moral
Foundations Theory research. It should be noted, however, the effect
size was small and the relationship was not found to be statistically
significant.

Upon speculation, the researchers identified one possible reason for why
the results were unable to confirm the hypothesis. The selection of the
broad and amorphous topic of \enquote{political news} may have led to
the scraping of large numbers of articles with little to no
moral-centric content. Rather, many articles may have been, for example,
simple reporting on congressional procedures that would leave little
room for the use of moral language here, let alone words from the Moral
Foundations Dictionary. In short, the range of topics covered in
Experiment 1 was likely too broad. The possibility exists that a tighter
focus on one political issue or event, especially one that (on the
surface) has a stronger relationship with morality might be more
illuminating for research in moral language in news media.

Owing to the exploratory nature of Experiment 1, the researchers were
afforded the opportunity to consider changes to the method to be
utilized in Experiment 2. Generally speaking, the researchers believe
their methodology to be sound. Web scraping methods and text processing
remain viable methods for collecting large amounts of text and
subsequently rendering that text in a form suitable for data analysis.
Experiment 1 also demonstrated a method by which to address inherent
problems in the Moral Foundations Dictionary relating to valence. The
solution provided in Experiment 1 appears to provide insights into the
MFD words where none previously existed. Finally, calculating weighted
percentages and sums for each moral foundation provides an easily
interpreted summary of MFD word positivity and occurrence.

While the methodology used in Experiment 1 features many strengths,
there are aspects which could be strengthened for future studies. The
researchers identified two such changes that were subsequently employed
in Experiment 2. First, the researchers elected to include more news
sources for web scraping and analysis in addition to the four used in
Experiment 1. Second, the researchers chose to focus their data
collection efforts exclusively on one event in U.S. politics: the
nomination and confirmation of Justice Brett Kavanaugh to the U.S.
Supreme Court. In Experiment 2, the researchers sought to confirm the
usefulness and validity of the method as well as test a similar
hypothesis as Experiment 1.

\section{Experiment 2}\label{experiment-2}

\subsection{Kavanaugh Supreme Court
Hearing}\label{kavanaugh-supreme-court-hearing}

In the wake of Justice Anthony Kennedy's retirement from the Supreme
Court of the United States, President Donald Trump nominated Brett
Kavanaugh as the new Associate Justice. Kavanaugh was previously on the
U.S. Court of Appeals for the District of Columbia. The Senate Judiciary
Committee began his confirmation hearing on September 4, 2018 (US
Government, 2018a). Following allegations of sexual assault by high
school classmate Dr.~Christine Blasey Ford, the committee postponed its
vote on whether or not to open the confirmation to the entire Senate.

On September 27, the committee questioned Dr.~Ford before commencing a
second round of questioning for Judge Kavanaugh (US Government, 2018b).
During the intervening weeks between hearings, two more women came
forward with two separate allegations of sexual assault on the part of
Kavanaugh. According to Nielsen reports, more than 20 million people
watched the September 27 proceedings on television (O'Connell, 2018).
This figure does not take into account viewers who watched online, nor
does it account for viewers outside the United States. On September 28,
the Senate Judiciary Committee voted to send the nomination to the
Senate floor. Senator Jeff Flake of Arizona, however, lobbied for a
week-long FBI investigation on Kavanaugh and the allegations facing him,
which the committee, and later the President, approved. The
investigation concluded with no significant findings. The Senate voted
50-48 to approve Kavanaugh's appointment on October 6, 2018 (US
Government, 2018c).

The Kavanaugh nomination, confirmation hearing, and eventual
swearing-in, as well as the news media's coverage of all three events,
feature many dimensions that likely differ depending on one's morals.
The issue might be exacerbated given the presence of questionable sexual
behaviors at the center of many concerns. On one side of the debate,
Kavanaugh's Supreme Court tenure presents a prime opportunity to bring
morality back into interpretation of the Constitution. Kavanaugh's
confirmation creates a conservative stronghold among the justices on the
court. Commentators have noted this might help advance a judicial agenda
that backpedals certain rights previously upheld by the Supreme Court,
including abortion and gay marriage - social issues challenged by their
opponents at least partially on moral grounds. Concerns around abortion
might be related either to \emph{harm/care} or \emph{purity/sanctity}.
On the other side of the debate, the assault allegations have energized
Kavanaugh's opponents to advocate for his rejection from the court owing
to misdeeds resulting from Kavanaugh's own alleged lack of morals.
Likewise, arguments could be made that relate concerns regarding sexual
violence to \emph{harm/care} or \emph{fairness/reciprocity}.
Additionally, the moral duty of the Senate as the upper chamber in the
U.S. legislature has been scrutinized in public discourse with respect
to its handling of the assault allegations vis-a-vis Kavanaugh's
confirmation.

\subsection{U.S. Government Shutdown of
2018-2019}\label{u.s.-government-shutdown-of-2018-2019}

The U.S. Federal Government partially shut down on December 22, 2018.
The government reopened on January 25, 2019 upon the passage of an
appropriations bill by both houses of Congress (Axelrod, 2019). The
shutdown stemmed from a disagreement between President Donald Trump and
Congress over funding for the president's proposed U.S.-Mexico border
wall. President Trump demanded \$5.7 billion dollars in the new budget
to be appropriated for the wall, which Congress did not provide in its
budget (Peoples \& Swanson, 2019). Owing to the lack of funding, the
government began a partial shutdown, causing \enquote{around 800,000}
federal employees to be either furloughed or compelled to work without
pay (Axelrod, 2019).

The government shutdown ended 35 days later when Congress passed a
continuing resolution to fund the government for three weeks while
further negotiations regarding the budget would take place. President
Trump signed the bill, ending the longest government shutdown in U.S.
history (Axelrod, 2019). Many political commentators cite the mounting
pressure on Trump and Congress stemming from significant costs both to
unpaid American workers and the economy as a whole (Kheel \& Mitchell,
2019). Perhaps owing to the financial pressures experienced by thousands
of federal employees and their families as a result of a political
quarrel, President Trump saw his approval rating fall as the shutdown
progressed (Peoples \& Swanson, 2019). While the shutdown and its
associated burdens were unpopular, the issues discussed concurrent to
the shutdown offer the potential for more divergent opinions.

As mentioned before, the shutdown started as a result of an impasse
regarding funding for the border wall. Since the wall was proposed
during the 2016 presidential campaign, the issue has taken on a moral
dimension. As an example, Shaun Casey, the director of Georgetown
University's Berkely Center for Religion, Peace and World Affairs, cites
diverging opinions stemming from theological concerns (Martin, 2019).
Some supporters, citing certain passages from the Bible, believe
construction of the wall is part of a divinely ordained process handed
down by God. Opponents, on the other hand, have cited other parts of the
Bible extolling the significance of exiled peoples, especially the
Hebrew people (Martin, 2019). This example potentially ties into the
five moral foundations. Theological concerns regarding the border wall
might invoke \emph{authority/respect} as well as \emph{purity/sanctity}.
In addition to these two foundations, the involvement of foreign
countries and their citizens' migration into the U.S. could also elicit
moral concerns related to \emph{ingroup/outgroup}.

Owing to the wealth of moral opinions regarding the border wall and its
association with the 2018-2019 government shutdown, the researchers
decided to add the government shutdown as a news event to analyze under
the auspices of Moral Foundations Theory. In addition to articles
related to the Kavanaugh confirmation hearing, the researchers scraped
articles related to the government shutdown in order to analyze their
content for valence and moral alignment.

\section{Method}\label{method-1}

In contrast to Experiment 1, the researchers approached Experiment 2
with the intention to confirm the method employed was valid for the
analysis of the scraped text as well as for any inferences drawn from
the analyses. For Experiment 2, the researchers hypothesized that news
sources perceived as liberal will exhibit positive endorsements of the
individualizing moral foundations (\emph{harm/care} and
\emph{fairness/reciprocity}) in their articles reporting on both the
Kavanaugh confirmation hearing as well as the 2018-2019 government
shutdown. News sources perceived as conservative are hypothesized to
positively endorse all five foundations equally in their coverage of the
Kavanaugh hearing and the government shutdown. The researchers tested
the hypothesis by analyzing the content scraped from news sources' web
pages spanning the two weeks before (September 13, 2018) and two weeks
after (October 11, 2018) Kavanaugh's confirmation hearing, owing to its
prominence in the news. Likewise, the researchers analyzed content
spanning two weeks before the start of the government shutdown (December
8, 2018) to two weeks following the end of the shutdown (February 8,
2019). The content will be analyzed for valence and moral alignment
under Moral Foundations Theory. Once again, no human participants were
needed for this study, so no Institutional Review Board approval was
necessary.

\subsection{Sources}\label{sources-1}

Articles pertaining to the Brett Kavanaugh Supreme Court confirmation
hearing and the 2018-2019 U.S. Government shutdown were scraped from the
websites of 10 U.S. news sources. As in Experiment 1, these sources were
selected owing to their favorability among political partisans according
to Mitchell et al. (2014). The sources favored by the highest proportion
of consistent liberals were \emph{The New York Times}, \emph{National
Public Radio (NPR)}, \emph{Slate}, \emph{Huffington Post}, and
\emph{Politico} (Mitchell et al., 2014). The sources favored by the
highest proportion of consistent conservatives included \emph{Fox News},
\emph{Breitbart}, \emph{The Rush Limbaugh Show}, \emph{The Blaze}, and
\emph{Sean Hannity}. Political articles referencing Brett Kavanaugh's
nomination process were identified and subsequently scraped by including
the URL for each source's coverage of the nomination in the \emph{R}
script. All code for this manuscript can be found at
\url{https://osf.io/5kpj7/}, and the scripts, again written with the
\emph{papaja} library in \emph{R}, are provided inline with this
manuscript (Aust \& Barth, 2017).

\subsection{Materials}\label{materials-1}

Using the \emph{rvest} library in the statistical package \emph{R}, we
pulled body text for individual articles from each of the aforementioned
10 news sources (identified using CSS language). We compiled the
articles into a dataset (Wickham, 2016). Using this dataset, we
identified word count and average word count per source. This process
was run for articles pertaining to Kavanaugh's nomination that were
published between September 13, 2018 and October 11, 2018 inclusive.
This date range was selected in reference to the widely-publicized and
viewed nomination hearing on September 27, 2018. We set the start date
at September 13 (two weeks before the hearing) and the end date at
October 11 (two weeks after the hearing) so that we could capture a
large amount of data (roughly one month) during which Kavanaugh's
nomination was at its peak saturation in news coverage.

The same process was followed for scraping articles related to the
partial U.S. Government shutdown of 2018-2019. The articles scraped were
published between December 8, 2018 and February 8, 2019 inclusive. Once
again, the researchers elected to scrape articles published two weeks
before and after the event in question in order to capitalize on the
shutdown's saturation in American news media.

\subsection{Data analysis}\label{data-analysis-1}

As in Experiment 1, the text was scanned with \emph{ngram}. Again, blank
articles, text from the Disqus system, and duplicate articles were
removed (Schmidt et al., 2017). The text was processed and stemmed in
order to convert to a usable form for further analysis (Feinerer \&
Hornik, 2017). Words were subsequently matched with their valences from
Warriner et al. (2013). Depending on the results of Jordan et al.
(2019)'s multi-trait multi-method analysis of the MFD and the Moral
Foundations Questionnaire, alternative forms of the Moral Foundations
Dictionary with additional words may be imported instead of the original
dictionary.

Using the \emph{tm} and \emph{ngram} packages in \emph{R}, the
researchers processed the text in order to convert it to lowercase, fix
spacing anomalies, and remove punctuation (Feinerer \& Hornik, 2017).
Each individual word was reduced to its stem (i.e., \emph{diseased} was
stemmed to \emph{diseas}). Once again, the same procedure was applied to
the MFD words and the words in the Warriner et al. (2013) dataset. Using
the Warriner et al. (2013) dictionary, the words in the MFD were
assigned their respective valence. The researchers obtained the words'
percent occurrence in the text. Percents were multiplied by
\emph{z}-scored valence and categorized into their proper MFD category.

\section{Results}\label{results-1}

\subsection{Descriptive Statistics}\label{descriptive-statistics-1}

The researchers calculated descriptive statistics for each news source
per topic in order to reveal the presence (if any) of linguistic
differences in the sources' use of language. As in Experiment 1,
statistics calculated include \emph{z}-scored valence, number of
articles per source, total words per source, mean tokens per article in
each source, mean types per article in each source, and mean readability
level (using the Flesch-Kincaid Grade Level Readability formula) per
source (Kincaid et al., 1975).

Table \ref{tab:exp2-source-descriptives-kav} displays the descriptive
statistics for sources' writing on the Kavanaugh confirmation hearing.
The sources were similar in most basic linguistic aspects, except for
number of articles. For example, \emph{Sean hannity} appears to have
published only 27 articles while \emph{Breitbart} published 757 articles
on this topic. Valence was found to be slightly positive across all
sources. \emph{Fox News} produced the most total words with the most
tokens on average. \emph{Politico} featured the highest number of types
on average. \emph{Rush Limbaugh} was the most readable source on average
by grade level while \emph{Slate} was the least readable. The large
standard deviations for these statistics, however, preclude conclusions
regarding differences in the sources' use of language.

Table \ref{tab:exp2-source-descriptives-gs} displays descriptive
statistics for articles about the partial government shutdown. Like the
Kavanaugh hearing, the sources were similar in average valence (slightly
positive). Once again, there was variation in the number of articles
published by each source on this topic. \emph{Sean Hannity}, \emph{Rush
Limbaugh}, and \emph{The Blaze} published fewer than 100 articles while
\emph{Fox News} published 1,013 articles. \emph{Fox News} again featured
the most total words and mean tokens. \emph{Politico} had the most types
on average. For this topic, \emph{Fox News} was the most readable while
\emph{Slate} was the least readable. For each statistic, the excessively
high standard deviations render any assertions regarding linguistic
differences inconclusive on a descriptive level.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp2-source-descriptives-kav}Kavanaugh - Descriptive Statistics by Source}
\small{
\begin{tabular}{lcccccclccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
Huffington Post & 0.27 & 0.12 & 552 & 359046 & 650.45 & 462.21 & 283.07 & 129.23 & 10.79 & 1.91\\
NPR & 0.22 & 0.25 & 366 & 108605 & 296.73 & 499.10 & 128.12 & 172.51 & 12.49 & 3.62\\
New York Times & 0.33 & 0.13 & 653 & 723569 & 1,108.07 & 570.31 & 461.26 & 174.20 & 10.56 & 2.10\\
Politico & 0.32 & 0.10 & 689 & 1069292 & 1,551.95 & 1,045.47 & 614.66 & 358.68 & 12.09 & 2.60\\
Slate & 0.29 & 0.12 & 272 & 229896 & 845.21 & 612.22 & 332.42 & 168.17 & 12.35 & 2.47\\
The Blaze & 0.25 & 0.13 & 277 & 128097 & 462.44 & 149.66 & 210.52 & 53.14 & 10.83 & 1.87\\
Breitbart & 0.30 & 0.13 & 757 & 375848 & 496.50 & 609.91 & 230.03 & 153.62 & 11.01 & 2.12\\
Fox News & 0.29 & 0.11 & 646 & 1304048 & 2,018.65 & 2,709.77 & 534.47 & 404.21 & 9.70 & 1.85\\
Sean Hannity & 0.31 & 0.11 & 27 & 5926 & 219.48 & 145.98 & 121.04 & 57.26 & 11.88 & 2.37\\
Rush Limbaugh & 0.38 & 0.13 & 172 & 267067 & 1,552.72 & 1,258.51 & 419.00 & 225.80 & 9.42 & 9.26\\
\bottomrule
\addlinespace
\end{tabular}
}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Readability statistics were calculated using the Flesch-Kincaid Grade Level readability formula. V = Valence, T = Tokens or total words, Ty = Types or unique words, FK = Flesch-Kincaid}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp2-source-descriptives-gs}Government Shutdown - Descriptive Statistics by Source}
\small{
\begin{tabular}{lcccccclccc}
\toprule
Source & $M_V$ & $SD_V$ & $N_{Article}$ & $N_{Words}$ & $M_T$ & $SD_T$ & $M_{Ty}$ & $SD_{Ty}$ & $M_{FK}$ & $SD_{FK}$\\
\midrule
Huffington Post & 0.30 & 0.13 & 432 & 242220 & 560.69 & 425.52 & 258.27 & 130.54 & 10.67 & 1.83\\
NPR & 0.23 & 0.22 & 434 & 151114 & 348.19 & 457.34 & 155.53 & 171.30 & 11.19 & 3.15\\
New York Times & 0.33 & 0.13 & 751 & 831184 & 1,106.77 & 506.92 & 471.80 & 174.88 & 11.28 & 2.19\\
Politico & 0.29 & 0.10 & 222 & 349499 & 1,574.32 & 875.63 & 638.32 & 313.34 & 11.28 & 1.23\\
Slate & 0.30 & 0.12 & 117 & 85254 & 728.67 & 464.34 & 301.71 & 138.74 & 11.86 & 2.51\\
The Blaze & 0.26 & 0.13 & 98 & 38983 & 397.79 & 102.10 & 185.66 & 41.93 & 10.68 & 1.85\\
Breitbart & 0.34 & 0.15 & 309 & 102886 & 332.96 & 238.70 & 168.31 & 81.59 & 10.71 & 2.05\\
Fox News & 0.35 & 0.13 & 1013 & 2799217 & 2,763.29 & 3,405.44 & 637.30 & 494.00 & 9.34 & 1.91\\
Sean Hannity & 0.29 & 0.14 & 63 & 10311 & 163.67 & 33.59 & 95.13 & 16.59 & 13.41 & 4.79\\
Rush Limbaugh & 0.37 & 0.12 & 78 & 152630 & 1,956.79 & 1,544.14 & 482.79 & 252.79 & 9.76 & 7.60\\
\bottomrule
\addlinespace
\end{tabular}
}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Readability statistics were calculated using the Flesch-Kincaid Grade Level readability formula. V = Valence, T = Tokens or total words, Ty = Types or unique words, FK = Flesch-Kincaid}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\subsection{Inferential Statistics}\label{inferential-statistics-1}

To analyze if news sources adhered to differences in word use based on
their target audience, the researchers utilized a multilevel model (MLM)
to analyze if news sources leveraged different vocabularies based on
target audience. The researchers used the \emph{nlme} library in
\emph{R} (Pinheiro et al., 2017). Each foundation's weighted percentage
was predicted by the source's political lean, using the individual
source as a random intercept to control for the nested data structure.
Two separate MLMs were constructed from datasets compiled for each topic
of interest: the Kavanaugh hearing and the partial government shutdown
of 2018-2019.

For the Kavanaugh topic, the multilevel model indicated the presence of
a statistically significant effect for \emph{harm/care}, but the
practical effect denoted by Cohen's \emph{d} was found to be small.
There were no other significant or practical effects of political lean
for any of the other four moral foundations. The effect for
\emph{harm/care} was in the hypothesized direction with liberal sources
tending to use more positively \emph{harm/care} words than conservative
sources. Descriptive and test statistics, \emph{p}-values and effect
sizes (Cohen's \emph{d}) can be found in Table \ref{tab:exp2-tablekav}.

For news articles about the partial U.S. Federal Government Shutdown of
2018-2019, there were no significant or practical effects of political
lean for the moral foundations. A small-to-medium effect size was
observed for \emph{authority/respect}. The effect was in the predicted
direction as well, as conservative sources tended to offer more positive
endorsements of the foundation than liberal sources. This effect was
similar to Experiment 1 in which the largest effect size was observed
for \emph{authority/respect}. As noted before, the effect found in
Experiment 1 was in the opposite direction as what was hypothesized.
Thus, it is difficult to draw comparisons between the two studies
despite the similar pattern for effect size. Owing to a lack of similar
effects for either Experiment 1 or the Kavanaugh topic, there is doubt
as to whether or not a practical or generalized effect exists for
\emph{authority/respect}.

Based on the weighted percent values for the five foundations applied to
both topics, MFD words seem to make up little of the article text. A
similar pattern was observed for the results in Experiment 1. As in
Experiment 1, the percentages and means seem to indicate a generally
positive endorsement of all five moral foundations across both political
leanings.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp2-tablekav}}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.39 & 1.35 & 0.66 & 1.35 & 2.89 & .020 & -0.19\\
Fairness/Reciprocity & 1.26 & 1.27 & 1.06 & 1.27 & -1.58 & .154 & 0.16\\
Ingroup/Loyalty & 1.15 & 1.22 & 1.06 & 1.22 & -0.63 & .549 & 0.07\\
Authority/Respect & 0.85 & 0.99 & 0.75 & 0.99 & -0.93 & .380 & 0.10\\
Purity/Sanctity & 0.90 & 1.28 & 1.06 & 1.28 & 0.88 & .406 & -0.11\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:exp2-tablegs}}
\begin{tabular}{lccccccc}
\toprule
Foundation & $M_C$ & $SD_C$ & $M_L$ & $SD_L$ & $t$ & $p$ & $d$\\
\midrule
Harm/Care & 0.91 & 1.27 & 1.01 & 1.27 & 1.68 & .132 & -0.07\\
Fairness/Reciprocity & 1.22 & 1.05 & 0.91 & 1.05 & -0.76 & .468 & 0.27\\
Ingroup/Loyalty & 1.38 & 1.16 & 1.11 & 1.16 & -0.48 & .644 & 0.21\\
Authority/Respect & 0.87 & 1.24 & 0.36 & 1.24 & -1.38 & .203 & 0.37\\
Purity/Sanctity & 1.29 & 1.13 & 0.99 & 1.13 & 0.03 & .979 & 0.22\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} For mean and standard deviation values, 'C' and 'L' refer to 'conservative' and 'liberal,' respectively}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\section{Discussion}\label{discussion-1}

The results obtained in Experiment 2 did not confirm either of the
hypotheses. The researchers found little compelling evidence of an
effect of partisan lean on MFD endorsement. The only significant effect
was found for \emph{harm/care} in the Kavanaugh confirmation hearing
dataset. While the effect was in the hypothesized direction (higher
positive endorsement for liberal sources), the effect size was small.
While there were no significant effects found for the government
shutdown topic, a small-to-medium effect size was calculated for the
\emph{authority/respect} foundation. Once again, the effect was in the
hypothesized direction with conservative sources providing higher
positive endorsements of the foundation relative to liberal sources.

Both of these effects are consistent with both the research hypotheses
as well as previous findings in Moral Foundations Theory scholarship. It
should be noted these effects were found in isolation, and similar
patterns were not found across the other studies. These findings
undermine any inferences of political bias stemming from the content of
news articles. The results obtained for Experiment 2 call into question
the Moral Foundations Dictionary's efficacy as a tool for establishing
differences between political news sources by partisan lean.

\section{Conclusions}\label{conclusions}

Within the theoretical framework of Moral Foundations Theory (Haidt \&
Graham, 2007), the researchers attempted to devise a method leveraging
the Moral Foundations Dictionary (Graham et al., 2009) in order to
quantify political bias stemming from content published by several
prominent American news sources. The researchers downloaded news article
text from the web, processed the text, matched it with the Moral
Foundations Dictionary words, and calculated average valence for words
in each article. To analyze the data, the researchers constructed
multi-level models for each foundation in which weighted percentages
were predicted by each source's political lean. Individual source was
held as a random intercept to control for correlated error. This process
was undertaken once in Experiment 1 for general political news and twice
in Experiment 2 using articles pertaining to two specific topics in
American news: Brett Kavanaugh's confirmation to the Supreme Court and
the most recent (as of this writing) partial U.S. Government shutdown
from December 2018 to January 2019.

In Experiment 1, the researchers analyzed general political news
articles from four prominent U.S. news sources: \emph{The New York
Times}, \emph{NPR}, \emph{Fox News}, and \emph{Breitbart}. They
hypothesized \emph{The New York Times} and \emph{NPR} would endorse
moral foundations representative of liberal political views and that
\emph{Fox News} and \emph{Breitbart} would endorse foundations
indicative of conservative views. The researchers assigned a political
lean (liberal or conservative) to the sources based upon consumer
preferences established by Mitchell et al. (2014). After collecting the
articles through web scraping, processing the article text, calculating
percent occurrence of MFD words, and weighting that value by valence,
the researchers constructed the MLM. The results obtained were not
significant in any statistical or practical sense. There was a small
effect found for the \emph{authority/respect} foundation according to
Cohen's \emph{d}, but the effect suggested endorsement for that
foundation opposite to what was hypothesized. The researchers viewed
these results as an opportunity to alter their method for use in
Experiment 2.

In Experiment 2, the researchers adjusted their method in order to
increase their potential ability to quantify political leanings in U.S.
news sources. First, the researchers expanded the number of news sources
from which articles were collected. They added \emph{Slate},
\emph{Politico}, and \emph{Huffington Post} to the list of liberal
sources and \emph{The Blaze}, \emph{Rush Limbaugh}, and \emph{Sean
Hannity} to the conservative sources. Once again, political bias groups
were assigned based on Mitchell et al. (2014)'s findings. Second, they
elected to collect articles and compile datasets pertaining to two
specific topics in U.S. news: the Kavanaugh hearing and the government
shutdown. These topics were chosen due to their relevance to multiple
moral foundations. Other than those two alterations, the same method
(scraping, processing, weighting, and multi-level modelling) was
employed for Experiment 2.

For articles pertaining to the Kavanaugh hearing, there was a
statistically significant effect for \emph{harm/care}, but the effect
size was modest. No other statistically or practically significant
effects were found for the Kavanaugh topic. Likewise, there were no
statistically or practically significant effects revealed for the
government shutdown topic. However, there was a small-to-medium effect
size for \emph{authority/respect}. The fact these patterns do not appear
in any of the other studies undertaken herein decreases the researchers'
ability to draw any valuable conclusions regarding the MFD's ability to
reveal political bias in American news media.

The results obtained in both experiments, while not confirmatory,
provide several avenues for future investigation. Twice confronted with
the failure to uncover differences in news media's political content,
the researchers were unable to confirm the existence of an effect of
political lean on American news writing under the auspices of Moral
Foundations Theory. This does not, however, rule out the possibility
that news media is biased. As Mitchell et al. (2014) has demonstrated,
self-reported political partisans are drawn to separate sources, but the
present research was unable to offer any content-based explanations of
this phenomenon. Even when attenuating the scraping criteria to collect
text concerning (what the researchers believed to be) morally resonant
topics, the method failed to detect differences in MFD word use by
prominent news sources. The researchers' focus, therefore, turns toward
the limitations of the method, especially the chosen theoretical
framework, under which the present research was conducted.

Despite the fact the results regarding political bias were inconclusive,
the researchers still retain confidence in the overall structure of the
methodology established in the current study. Specifically, the
procedure for scraping text from the web, processing, stemming, and
weighting the scores with valence seems to represent a solid method for
preparing a high quantity of text passages for data analysis. The
researchers implemented valence as an indicator of the directionality of
endorsement due to the inherent ambiguity of simply calculating MFD word
percent occurrence. This method served both to augment the face validity
of the MFD by incorporating valence (thus reducing ambiguity) and to
generate a score that is easy to understand and analyze.

Thus, the researchers' attention turned to the instrument used to detect
differences. In the present study, the researchers utilized the Moral
Foundations Dictionary, a metric developed from Moral Foundations Theory
(Graham et al., 2009). Based upon the results obtained, it might be
necessary to investigate alternative instruments that might better
elucidate the differences of interest. Likewise, other theoretical
perspectives may be better equipped to explain political differences in
discourse.

Tools such as these could be developed through the Linguistic Inquiry
and Word Count program (LIWC) (Pennebaker et al., 2007). As stated
before, the Moral Foundations Dictionary was established through an
analysis of Christian sermons performed mostly with LIWC as well as some
human-enforced corrections due to context (Graham et al., 2009). This
approach was grounded in Moral Foundations Theory, a clear framework
within which the authors developed the MFD. The original authors of the
MFD established sound psychometric properties for the tool, implying its
validity in several applications. For this reason, the researchers
elected to utilize the MFD as their primary instrument.

Other instruments could theoretically be developed through a LIWC
analysis, including those that would be useful for analyzing political
discourse. Researchers could obtain percent occurrence scores for words
appearing in certain LIWC categories, which could then be weighted with
valence scores from Warriner et al. (2013). The researchers will
consider other applications of the LIWC in future studies aiming to
analyze political content. James Pennebaker (2007), one of the creators
of LIWC, has suggested the program could be applied to analyses of
cognitive, social, and emotional states. Furthermore, there is some
evidence LIWC can be used to detect deception, but the reliability of
the program's accuracy in this regard (67\%) calls this assertion into
question (Pennebaker, 2011). Despite the promising extent of LIWC's
abilities, without a firm theoretical foundation for such analyses, any
such applications of LIWC would face serious questions as to their
relevance, validity, and generalizability. Nevertheless, LIWC analyses
may represent a path to further research into political differences in
news media as well as theory formation.

There is, however, another theoretical framework that could prove useful
when applied to the method established in the present research.
Relational Frame Theory originates in applied behavior analysis, but
there exists recent scholarship that demonstrates its potential
usefulness in the political realm. Relational Frame Theory (RFT) holds
that stimuli, including speech and discourse, elicit responses in terms
of relationships to other stimuli, even when the stimuli do not
otherwise appear to be related (Blackledge, 2014). For example, an
individual's equivalent fear response to a venomous snake as well as a
wooded area said to \enquote{contain snakes} represents \enquote{mutual
entailment.} That is, the fear response is an effect of both the snake
and environment stimuli, which cause the fear response.
\enquote{Combinatorial entailment} can be represented by a fear response
elicited by a novel stimulus whose relationship to a well-known stimulus
elicits a certain response (Blackledge, 2014). In the snake example, if
another stimulus (Stimulus A), were said to be \enquote{worse than} the
snake, the relationship between A and the snake would elicit a more
severe fear response to Stimulus A. RFT utilizes a set of
\enquote{families} describing the nature of these relationships. Such
families include \enquote{coordinative,} establishing similarities
between stimuli, and \enquote{distinctive,} implying fundamental
differences between stimuli (Belisle, Paliliunas, Dixon, \& Tarbox,
2018). These families can be used to analyze networks of relational
frames in discourse.

Belisle et al. (2018) conducted a text analysis (using a computer
algorithm) of several speeches made by four recent U.S. presidents under
Relational Frame Theory. These speeches, representing prominent examples
of political discourse, were found to differ in their use of relational
frames. For example, speeches by Donald Trump were found to contain more
distinctive frames than other presidents, while Barack Obama utilized
more coordinative frames than his counterparts (Belisle et al., 2018).
At first glance, the differences appear to fall along party lines,
though other factors may be involved as well. In any case, this research
appears to indicate that RFT might be useful for analyzing political
news content similar to the method employed in the current study.

Another apparent strength of RFT is its demonstrated effectiveness in
the analysis of political discourse. Unlike the development of the Moral
Foundations Dictionary, this application of RFT was performed on public
facing political discourse that might be more similar to news articles
than the Christian sermons upon which the MFD was compiled.
Additionally, there would likely be greater overlap between the speeches
analyzed in Belisle et al. (2018)'s work and political news, as the
media often quotes speeches by the president. Furthermore, the types or
\enquote{families} of relational frames could be used as the individual
constructs much like the moral foundations in the present research. In
order to conform with the method established for this study, their
occurrence would need to be scored, probably as a percent. This would
allow for an intuitive approach to weighting by \emph{z}-scored valence
if needed.

While Relational Frame Theory appears to represent a viable option for
further research into political news analysis, there are some drawbacks.
First, RFT by itself does not purport to elucidate political differences
in discourse, and its applications to political research are not yet
well established. In other words, RFT represents one method by which to
analyze political text, but with the caveat that the categories are not
known to represent patterns of political ideology. One of the main
strengths of Moral Foundations Theory in the eyes of the researchers was
that it was a theory designed to investigate not only moral differences
but also political differences. While the results obtained herein may
weaken assumptions regarding the MFD's capabilities with political text,
Moral Foundations Theory has been repeatedly established as a strong
framework for detecting political differences along moral lines.

In order to seriously consider RFT as a viable framework for this type
of analysis, the researchers would like to be able to draw on further
confirmatory evidence of the theory's utility in political discourse
analysis. Specifically, analyses of more individuals than the four
presidents addressed by Belisle et al. (2018) would be needed to further
demonstrate RFT's applicability in this field. Also, though presidential
speeches may be more similar to news articles than sermons, they are
still a distinct form of discourse with different goals and audiences
than news outlets. RFT's reliability across these forms appears to
remain untested as of this writing. These potential shortcomings are
testable and might be overcome through more research. As such, RFT's
potential as a useful framework for political discourse analysis should
not be discounted.

The researchers aimed to utilize the Moral Foundations Dictionary and
valence in the analysis of political news articles to determine whether
or not bias was present. In Experiment 1, they tested general political
news from four U.S. news sources - two liberal and two conservative.
They were unable to find any significant evidence of political bias in
that analysis. After adjusting the method to include more sources and
specific topics for articles, including the Kavanaugh hearing and the
government shutdown, the researchers were once again unable to uncover
differences in MFD word use between conservative and liberal news
sources. The results in the current study seem to demonstrate the MFD's
weaknesses in applications to political discourse. MFD words simply
might not appear often enough in news articles to register as
endorsement for any particular moral foundation over the others. In
future research, this weakness could be addressed through further
alterations to the method, including alternative analyses using computer
programs including LIWC, or through a different theoretical perspective
such as Relational Frame Theory.

Never before has news media and discourse represented such fertile
ground for psychological research. The plethora of options for news
sources has created not only an abundance of choice but also vast
quantities of text data. Along with this recent increase in the amount
of text information, there is now an obligation on the part of
researchers to devise proper methods for analyzing that text. Solid
methodologies must be constructed and periodically improved to keep pace
with evolving technologies. Likewise, a strong theoretical foundation is
paramount to making sense of the current and future media ecosystem.
Therefore, it is incumbent upon social scientists to continue
investigating the information consumed by millions of Americans every
day so that insights into the nature and consequences of political
discourse can be more completely understood.

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-Aust2017}{}
Aust, F., \& Barth, M. (2017). papaja: Create APA manuscripts with R
Markdown. Retrieved from \url{https://github.com/crsh/papaja}

\hypertarget{ref-Axelrod2019}{}
Axelrod, T. (2019). Poll: Majority of Americans hold Trump and
Republicans responsible for shutdown \textbar{} TheHill. Retrieved from
\url{https://thehill.com/homenews/campaign/427007-poll-majority-of-americans-hold-trump-and-republicans-responsible-for}

\hypertarget{ref-Belisle2018}{}
Belisle, J., Paliliunas, D., Dixon, M. R., \& Tarbox, J. (2018).
Feasibility of contextual behavioral speech analyses of US presidents:
Inaugural addresses of Bill Clinton, George W. Bush, Barack Obama, and
Donald Trump, 1993--2017. \emph{Journal of Contextual Behavioral
Science}, \emph{10}(July), 14--18.
doi:\href{https://doi.org/10.1016/j.jcbs.2018.07.002}{10.1016/j.jcbs.2018.07.002}

\hypertarget{ref-Blackledge2014}{}
Blackledge, J. T. (2014). An introduction to relational frame theory:
Basics and applications. \emph{The Behavior Analyst Today}, \emph{3}(4),
421--433. doi:\href{https://doi.org/10.1037/h0099997}{10.1037/h0099997}

\hypertarget{ref-Bradley1999}{}
Bradley, M. M., \& Lang, P. J. (1999). \emph{Affective Norms for English
Words (ANEW): Instruction manual and affective ratings} (No. C-1). The
Center for Research in Psychophysiology, University of Florida.

\hypertarget{ref-Cameron2006}{}
Cameron, D. (2006). Ideology and language. \emph{Journal of Political
Ideologies}, \emph{11}(2), 141--152.
doi:\href{https://doi.org/10.1080/13569310600687916}{10.1080/13569310600687916}

\hypertarget{ref-Clinton2014}{}
Clinton, J. D., \& Enamorado, T. (2014). The national news media's
effect on Congress: How Fox News affected elites in Congress. \emph{The
Journal of Politics}, \emph{76}(4), 928--943.
doi:\href{https://doi.org/10.1017/S0022381614000425}{10.1017/S0022381614000425}

\hypertarget{ref-Davies2014}{}
Davies, C. L., Sibley, C. G., \& Liu, J. H. (2014). Confirmatory factor
analysis of the moral foundations questionnaire independent scale
validation in a new zealand sample. \emph{Social Psychology},
\emph{45}(6), 431--436.
doi:\href{https://doi.org/10.1027/1864-9335/a000201}{10.1027/1864-9335/a000201}

\hypertarget{ref-Druckman2004}{}
Druckman, J. N., Jacobs, L. R., \& Ostermeier, E. (2004). Candidate
strategies to prime issues and image. \emph{The Journal of Politics},
\emph{66}(4), 1180--1202.
doi:\href{https://doi.org/10.1111/j.0022-3816.2004.00295.x}{10.1111/j.0022-3816.2004.00295.x}

\hypertarget{ref-Feinerer2017}{}
Feinerer, I., \& Hornik, K. (2017). Text mining package. Retrieved from
\url{http://tm.r-forge.r-project.org/}

\hypertarget{ref-Gelman2006}{}
Gelman, A. (2006). Multilevel (hierarchical) modeling: What it can and
cannot do. \emph{Technometrics}, \emph{48}(3), 432--435.
doi:\href{https://doi.org/10.1198/004017005000000661}{10.1198/004017005000000661}

\hypertarget{ref-Gilligan1982}{}
Gilligan, C. (1982). New maps of development: New visions of maturity.
\emph{American Journal of Orthopsychiatry}, \emph{52}(2), 199--212.
doi:\href{https://doi.org/10.1111/j.1939-0025.1982.tb02682.x}{10.1111/j.1939-0025.1982.tb02682.x}

\hypertarget{ref-Graham2009}{}
Graham, J., Haidt, J., \& Nosek, B. A. (2009). Liberals and
conservatives rely on different sets of moral foundations. \emph{Journal
of Personality and Social Psychology}, \emph{96}(5), 1029--1046.
doi:\href{https://doi.org/10.1037/a0015141}{10.1037/a0015141}

\hypertarget{ref-Graham2011}{}
Graham, J., Nosek, B. A., Haidt, J., Iyer, R., Koleva, S., \& Ditto, P.
H. (2011). Mapping the moral domain. \emph{Journal of Personality and
Social Psychology}, \emph{101}(2), 366--385.
doi:\href{https://doi.org/10.1037/a0021847}{10.1037/a0021847}

\hypertarget{ref-Haidt2007}{}
Haidt, J., \& Graham, J. (2007). When morality opposes justice:
Conservatives have moral intuitions that Liberals may not recognize.
\emph{Social Justice Research}, \emph{20}(1), 98--116.
doi:\href{https://doi.org/10.1007/s11211-007-0034-z}{10.1007/s11211-007-0034-z}

\hypertarget{ref-Jordan2019}{}
Jordan, K. N., Buchanan, E. M., \& Padfield, W. E. (2019). \emph{A
validation of the Moral Foundations Questionnaire and Dictionary}.
Retrieved from \url{https://osf.io/kt9yf/}

\hypertarget{ref-Kheel2019}{}
Kheel, R., \& Mitchell, E. (2019). Overnight Defense: Trump agrees to
reopen government without wall funding \textbar{} Senate approves
stopgap spending \textbar{} Dems ask Armed Services chair to block
military funding for wall \textbar{} Coast Guard official assures
workers they will receive back pay \textbar{} TheHill. Retrieved from
\url{https://thehill.com/policy/defense/overnights/427045-overnight-defense-trump-agrees-to-reopen-government-without-wall}

\hypertarget{ref-Kincaid1975}{}
Kincaid, J. P., Fishburne, J., Robert P., R., Richard L., C., \& Brad S.
(1975). Derivation of new readability formulas (Automated Readability
Index, Fog Count and Flesch Reading Ease Formula) for Navy enlisted
personnel.
doi:\href{https://doi.org/10.21236/ADA006655}{10.21236/ADA006655}

\hypertarget{ref-Kohlberg1977}{}
Kohlberg, L., \& Hersh, R. H. (1977). Moral development: A review of the
theory. \emph{Theory into Practice}, \emph{16}(2), 53--59.
doi:\href{https://doi.org/10.1080/00405847709542675}{10.1080/00405847709542675}

\hypertarget{ref-Kuperman2014}{}
Kuperman, V., Estes, Z., Brysbaert, M., \& Warriner, A. B. (2014).
Emotion and language: Valence and arousal affect word recognition.
\emph{Journal of Experimental Psychology: General}, \emph{143}(3),
1065--1081.
doi:\href{https://doi.org/10.1037/a0035669}{10.1037/a0035669}

\hypertarget{ref-Kuperman2012}{}
Kuperman, V., Stadthagen-Gonzalez, H., \& Brysbaert, M. (2012).
Age-of-acquisition ratings for 30,000 English words. \emph{Behavior
Research Methods}, \emph{44}(4), 978--990.
doi:\href{https://doi.org/10.3758/s13428-012-0210-4}{10.3758/s13428-012-0210-4}

\hypertarget{ref-Martin2019}{}
Martin, M. (2019). The moral question opf Trump's border wall. NPR.
Retrieved from
\url{https://www.npr.org/2019/01/27/689191255/the-morality-question-of-trump-s-border-wall}

\hypertarget{ref-Mitchell2014}{}
Mitchell, A., Matsa, K. E., Gottfried, J., \& Kiley, J. (2014).
Political polarization \& media habits \textbar{} Pew Research Center.
Retrieved from
\url{http://www.journalism.org/2014/10/21/political-polarization-media-habits/}

\hypertarget{ref-New2007}{}
New, B., Brysbaert, M., Veronis, J., \& Pallier, C. (2007). The use of
film subtitles to estimate word frequencies. \emph{Applied
Psycholinguistics}, \emph{28}(4), 661--677.
doi:\href{https://doi.org/10.1017/S014271640707035X}{10.1017/S014271640707035X}

\hypertarget{ref-OConnell2018}{}
O'Connell, M. (2018). Ford-Kavanaugh Ratings: Hearing brings 20 million
viewers to cable and broadcast \textbar{} Hollywood Reporter. Retrieved
from
\url{https://www.hollywoodreporter.com/live-feed/ford-kavanaugh-ratings-hearing-brings-20-million-viewers-cable-broadcast-1147785}

\hypertarget{ref-Pashler2012}{}
Pashler, H., Coburn, N., \& Harris, C. R. (2012). Priming of Social
Distance? Failure to Replicate Effects on Social and Food Judgments.
\emph{PLoS ONE}, \emph{7}(8).
doi:\href{https://doi.org/10.1371/journal.pone.0042510}{10.1371/journal.pone.0042510}

\hypertarget{ref-Pennebaker2011}{}
Pennebaker, J. W. (2011). Using computer analyses to identify language
style and aggressive intent: The secret life of function words.
\emph{Dynamics of Asymmetric Conflict}, \emph{4}(2), 92--102.
doi:\href{https://doi.org/10.1080/17467586.2011.627932}{10.1080/17467586.2011.627932}

\hypertarget{ref-Pennebaker2007}{}
Pennebaker, J. W., Booth, R. J., \& Frances, M. E. (2007). Liwc2007:
Linguistic inquiry and word count. Austin, TX.

\hypertarget{ref-Peoples2019}{}
Peoples, S., \& Swanson, E. (2019, January). Trump's approval rating
sinks in new poll as he gets most blame for shutdown. Chicago.

\hypertarget{ref-Pinheiro2017}{}
Pinheiro, J., Bates, D., Debroy, S., Sarkar, D., \& Team, R. C. (2017).
nlme: Linear and nonlinear mixed effects models. Retrieved from
\url{https://cran.r-project.org/package=nlme}

\hypertarget{ref-Schmidt2017}{}
Schmidt, M. F., Gonzalez-Cabrera, I., \& Tomasello, M. (2017).
Children's developing metaethical judgments. \emph{Journal of
Experimental Child Psychology}, \emph{164}, 163--177.
doi:\href{https://doi.org/10.1016/j.jecp.2017.07.008}{10.1016/j.jecp.2017.07.008}

\hypertarget{ref-Suhler2011}{}
Suhler, C. L., \& Churchland, P. (2011). Can innate, modular
``foundations'' explain morality? Challenges for Haidt's Moral
Foundations Theory. \emph{Journal of Cognitive Neuroscience},
\emph{23}(9), 2103--2116.
doi:\href{https://doi.org/10.1162/jocn.2011.21637}{10.1162/jocn.2011.21637}

\hypertarget{ref-USGovernment2018b}{}
US Government. (2018a). Congressional Record. \emph{Congressional
Record}, \emph{164}(146), 46.
doi:\href{https://doi.org/10.1097/00017285-197507000-00018}{10.1097/00017285-197507000-00018}

\hypertarget{ref-USGovernment2018}{}
US Government. (2018b). Congressional Record. \emph{Congressional
Record}, \emph{164}(160), 93.
doi:\href{https://doi.org/10.1097/00017285-197507000-00018}{10.1097/00017285-197507000-00018}

\hypertarget{ref-USGovernment2018a}{}
US Government. (2018c). Congressional Record. \emph{Congressional
Record}, \emph{164}(167), 14.
doi:\href{https://doi.org/10.1097/00017285-197507000-00018}{10.1097/00017285-197507000-00018}

\hypertarget{ref-Warriner2015}{}
Warriner, A. B., \& Kuperman, V. (2015). Affective biases in English are
bi-dimensional. \emph{Cognition and Emotion}, \emph{29}(7), 1147--1167.
doi:\href{https://doi.org/10.1080/02699931.2014.968098}{10.1080/02699931.2014.968098}

\hypertarget{ref-Warriner2013}{}
Warriner, A. B., Kuperman, V., \& Brysbaert, M. (2013). Norms of
valence, arousal, and dominance for 13,915 English lemmas.
\emph{Behavior Research Methods}, \emph{45}(4), 1191--1207.
doi:\href{https://doi.org/10.3758/s13428-012-0314-x}{10.3758/s13428-012-0314-x}

\hypertarget{ref-Wickham2016}{}
Wickham, H. (2016). Package ` rvest '. Retrieved from
\url{https://cran.r-project.org/package=rvest}

\hypertarget{ref-Williams2008}{}
Williams, L. E., \& Bargh, J. A. (2008). Keeping one's distance.
\emph{Psychological Science}, \emph{19}(3), 302--308.
doi:\href{https://doi.org/10.1111/j.1467-9280.2008.02084.x}{10.1111/j.1467-9280.2008.02084.x}

\endgroup

\end{document}