### 2020-08-05 manuscript/Partisan Media Manuscript.Rmd

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
#install.packages(c("devtools", "rio", "forestplot", "nlme", "MOTE", "tools"))
#devtools::install_github("crsh/papaja")
library(papaja) 
library(rio)
library(forestplot)
library(nlme)
library(MOTE)
library(tools)

#setwd('C:/Users/kayla/Documents/GitHub/Will-Pilot/manuscript')
nsim = 10

####make the table for study 1####
table1 = matrix(NA, ncol = 12, nrow = 23)
colnames(table1) = c("DV", "Mconservative", 
                     "SDconservative", "Mliberal", "SDliberal", 
                     "d", "dlow", "dhigh", "bcilowC", "bcihighC", 
                     "bcilowL", "bcihighL")
table1 = as.data.frame(table1)

####make the table for study 2: government shutdown####
table2 = matrix(NA, ncol = 12, nrow = 23)
colnames(table2) = c("DV", "Mconservative", 
                     "SDconservative", "Mliberal", "SDliberal", 
                     "d", "dlow", "dhigh", "bcilowC", "bcihighC", 
                     "bcilowL", "bcihighL")
table2 = as.data.frame(table2)

####make the table for study 2: Kavanaugh Hearings####
table3 = matrix(NA, ncol = 12, nrow = 23)
colnames(table3) = c("DV", "Mconservative", 
                     "SDconservative", "Mliberal", "SDliberal", 
                     "d", "dlow", "dhigh", "bcilowC", "bcihighC", 
                     "bcilowL", "bcihighL")
table3 = as.data.frame(table3)
```

```{r bootstrap-function, include = FALSE}
####bootstrap function #####
bootstrap_values <- function(formula, dataset, nsim){

  store_mean <- rep(NA, nsim)
  store_sd <- rep(NA, nsim)  
  attempts <- 0
  #loop until you have enough
  while(attempts < nsim){
    #create a dataset
    d <- dataset[sample(1:nrow(dataset), 
                        size = nrow(dataset),
                        replace = TRUE), ]
    #test the model
    tryCatch({

      model1 = lme(formula, 
             data = d, 
             method = "ML", 
             na.action = "na.omit",
             random = list(~1|Source),
             control=lmeControl(opt = "nlminb")) 
      meanvalue = summary(model1)$tTable[1]
      sdvalue = summary(model1)$tTable[2] * sqrt(nrow(d))
      attempts <- attempts + 1
      store_mean[attempts] <- meanvalue
      store_sd[attempts] <- sdvalue
      return(store_mean, store_sd, attempts)
    }, error = function(x){})
  }

  return(list("mean" = store_mean, "sd" = store_sd))
}

```

# Experiment 1

move the data collection from the original thesis

```{r bootstrap-study-1, include = F}
# Import the data
exp1 <- import("finaldata_LIWC.csv")

# Eliminate Missing Data
exp1 <- na.omit(exp1)

# Create group labels
exp1$PartisanLean = ifelse(exp1$PartisanLean=='Conservative', 0, 1)

# Create IVs and DVs
groups = c(0, 1)
DVs = c('Analytic','Clout','Authentic','Tone','WPS','Sixltr','anger','social','family','female','male','cogproc','affiliation','achieve','power','reward','risk','focuspast','focuspresent','focusfuture','money','relig','death')

# Loop over the DVs and create values 
# This saves the difference between groups
r = 1
for(DV in DVs){
  for(group in groups){
    data = subset(exp1, PartisanLean==group) #data name
    f = as.formula(paste(DV,'~1', sep=''))
    bs = bootstrap_values(f,data,nsim)
    table1[r, 1] = DV
    if(group==0){
      table1[r, 2] = mean(bs$mean) #table name all the way down 
      table1[r, 3] = mean(bs$sd)
      ncon = length(na.omit(data[[DV]]))
      table1[r, 9] = quantile(bs$mean, 0.025)
      table1[r, 10] = quantile(bs$mean, 0.975)
    }
    if(group==1){
      table1[r, 4] = mean(bs$mean)
      table1[r, 5] = mean(bs$sd)
      nlib = length(na.omit(data[[DV]]))
      table1[r, 11] = quantile(bs$mean, 0.025)
      table1[r, 12] = quantile(bs$mean, 0.975)
    }
  }
  mdiff = d.ind.t(m1 = table1[r,2], m2 = table1[r,4], 
                sd1 = table1[r,3], sd2 = table1[r,5],
                n1 = ncon, 
                n2 = nlib, 
                a = .05)
  table1[r, 6] = mdiff$d
  table1[r, 7] = mdiff$dlow
  table1[r, 8] = mdiff$dhigh
  r = r + 1
}
```

Doing the bootstrapping thing from the language of war paper
- analytic, cognitive processing, authenticity, tone, clout

- put notes here, I fixed the code

# Experiment 2

move the data collection from the original thesis
both effect sizes and topics analysis

```{r bootstrap-study-2-goverment-shutdown}
# Import the Data
exp2 <- import("exp2_merged data_LIWC.csv")
colnames(exp2)[1:5] = c('Source', 'URL', 'Text', 'Event', 'PartisanLean')

# Eliminate Missing Data

#subset with Event=='gs'
exp2_gs <- subset(exp2, Event == "gs")

```

```{r bootstrap-study-2-kavanaugh-hearings}

#subset with Event=='kav'
exp2_kav ... 
```

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup