### 2021-07-01 manuscript/topic_model.Rmd
here's the topic model

---
title: "Topic Model"
author: "Erin M. Buchanan"
date: "7/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(topicmodels)
library(rio)
library(tm)
library(slam)
library(ggplot2)
library(tidyverse)
library(tidytext)
library(dplyr)
```

## Data

```{r}
DF <- import("finaldata_LIWC.csv")
```

## Text Cleaning

```{r}
DF$clean <- removeNumbers(DF$Text)
DF$clean <- removePunctuation(DF$clean)
DF$clean <- removeWords(DF$clean, stopwords(kind = "SMART"))
DF$clean <- tolower(DF$clean)
DF$clean <- gsub("[[:punct:]]", " ", DF$clean)
DF$clean <- gsub("\n", " ", DF$clean)
```

## Topic Model

```{r}
import_corpus <- Corpus(VectorSource(DF$clean))
import_matrix <- DocumentTermMatrix(import_corpus)
import_weight <- tapply(import_matrix$v/row_sums(import_matrix)[import_matrix$i], 
                       import_matrix$j, 
                       mean) *
  log2(nDocs(import_matrix)/col_sums(import_matrix > 0))

import_matrix <- import_matrix[ row_sums(import_matrix) > 0, ]

k <- 10 
SEED <- 58934

LDA_fit <- LDA(import_matrix, k = k, 
              control = list(seed = SEED))
```

## Results

```{r}
terms(LDA_fit,20)

LDA_fit_topics <- tidy(LDA_fit, matrix = "beta")
top_terms <- LDA_fit_topics %>%
   group_by(topic) %>%
   top_n(15, beta) %>%
   ungroup() %>%
   arrange(topic, -beta)

top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  theme_classic() +
  coord_flip()
```