### 2024-11-30 manuscript_version/revision/appendix.Rmd
rearranging folders and updating the manuscript

---
title: "appendix.Rmd"
---

# Study 2: Descriptive Statistics

```{r exp1_descriptive_calc, echo = F, include = F}
final$uniquewords = NA
final$valenceavg = NA

for (i in 1:nrow(final)){

  #add unique
  final$uniquewords[i] = length(unique(unlist(strsplit(final$stemmed[i], " "))))

  #add valence
  tempwords = as.data.frame(unlist(strsplit(final$stemmed[i], " ")))
  colnames(tempwords) = "theword"
  tempwords$tempval = vlookup(tempwords$theword, #index in original data
                    warriner, #where's the stuff you want to put in
                    "V.Mean.Sum", #what column is the stuff you want
                    "Word2") #what's the matching index

  final$valenceavg[i] = mean(tempwords$tempval, na.rm = T)

}

# tapply(final$valenceavg, final$Source, mean)
# tapply(final$valenceavg, final$Source, sd)
# tapply(final$wordcount, final$Source, mean)
# tapply(final$wordcount, final$Source, sd)
# tapply(final$uniquewords, final$Source, mean)
# tapply(final$uniquewords, final$Source, sd)
# tapply(final$Source, final$Source, length)
# tapply(final$wordcount, final$Source, sum)

final$FK = NA

for (i in 1:nrow(final)){
  final$FK[i] = unlist(textstat_readability(final$Text[i],
                     measure = "Flesch.Kincaid")$Flesch.Kincaid)
}

# tapply(final$FK, final$Source, mean)
# tapply(final$FK, final$Source, sd)

####build the table here####
tableprint = matrix(NA, nrow = 4, ncol = 11)
colnames(tableprint) = c("Source", 
                         "Mean_Valence", "SD_Valence", "Number_of_Articles",
                         "Total_Words", "Mean_Tokens", "SD_Tokens", "Mean_Types",
                         "SD_Types", "Mean_Readability", "SD_Readability")

#tableprint[ , 1] = unique(final$Source)
tableprint[ , 1] = c("Breitbart", "Fox News", "NPR", "New York Times")
tableprint[ , 2] = tapply(final$valenceavg, final$Source, mean)
tableprint[ , 3] = tapply(final$valenceavg, final$Source, sd)
tableprint[ , 4] = tapply(final$valenceavg, final$Source, length)
tableprint[ , 5] = tapply(final$wordcount, final$Source, sum)
tableprint[ , 6] = tapply(final$wordcount, final$Source, mean)
tableprint[ , 7] = tapply(final$wordcount, final$Source, sd)
tableprint[ , 8] = tapply(final$uniquewords, final$Source, mean)
tableprint[ , 9] = tapply(final$uniquewords, final$Source, sd)
tableprint[ , 10] = tapply(final$FK, final$Source, mean)
tableprint[ , 11] = tapply(final$FK, final$Source, sd)

tableprint[ , c(2,3,6,7,8,9,10,11)] = printnum(as.numeric(tableprint[ , c(2,3,6,7,8,9,10,11)]), big.mark = "")

tableprint = tableprint[c(3, 4, 1, 2) , ]
```

We calculated descriptive statistics for each news source in order to
understand any and all fundamental linguistic differences in the
sources' use of English. Statistics calculated included average
*z*-scored valence of the unique words per article, number of articles
per source, total number of words per source, average number of tokens
(words) per article in each source, average number of types (unique
words) per article in each source, and mean readability level per
source. Readability statistics were calculated using the Flesch-Kincaid
Grade Level Readability formula [@Kincaid1975]. Readability is
calculated using a formula where the total number of syllables, words,
and sentences in a given passage are determinants of its difficulty. The
obtained value is intended to match up with the U.S. grade level at
which one should be able to comfortably read the passage [@Kincaid1975].
For example, a text with a readability score of 11 should be easily read
by a U.S. high school junior.

As seen in Table \@ref(tab:exp1-source-descriptives), the sources are
similar in some aspects yet different in others. Valence appears to be
slightly positive across all sources. The large standard deviations seem
to indicate little to no presence of a difference in valence across
sources. *NYT* published the greatest number of articles as well as
total words. *Breitbart* featured the lowest number of articles, and
*NPR* the lowest number of total words from all articles. Per individual
article, however, *Breitbart* appears to feature the highest average
number of words as well as unique words. Once again the standard
deviations call into question any apparent differences between sources.
Finally, *Fox News* articles had the lowest reading grade level on
average, while *NYT* had the highest. This result might be attributable
to the greater number of tokens in the average *NYT* article compared to
*Fox News*. The standard deviations for readability suggest the presence
of a diverse array of articles for each source, ranging from low to high
reading level. Large standard deviations suggest the sources feature a
lot of overlap between them in their representation of scores.

```{r exp1-source-descriptives, results = 'asis', echo = FALSE}
#change this to match
apa_table(as.data.frame(tableprint),
          note = "Readability statistics were calculated using the Flesch-Kincaid Grade Level readability formula. V = Valence, T = Tokens or total words, Ty = Types or unique words, FK = Flesch-Kincaid",
          caption = "Study 2 - Descriptive Statistics by Source",
          col.names = c("Source", "$M_V$", "$SD_V$", "$N_{Article}$", "$N_{Words}$", "$M_T$", "$SD_T$", "$M_{Ty}$", "$SD_{Ty}$", "$M_{FK}$", "$SD_{FK}$"),
          format = "latex",
          escape = F,
          align = c("l", rep("c", 10)),
          font_size = 'footnotesize',
          placement = "h"
          )
```

# Study 3: Descriptive Statistics

We calculated descriptive statistics for each news source per topic in
order to reveal the presence (if any) of linguistic differences in the
sources' use of language. As in Study 1, statistics calculated include
*z*-scored valence, number of articles per source, total words per
source, mean tokens per article in each source, mean types per article
in each source, and mean readability level (using the Flesch-Kincaid
Grade Level Readability formula) per source [@Kincaid1975].

Table \@ref(tab:exp2-source-descriptives-kav) displays the descriptive
statistics for sources' writing on the Kavanaughnaugh confirmation
hearing. The sources were similar in most basic linguistic aspects,
except for number of articles. For example, *Sean Hannity* appears to
have published only 27 articles while *Breitbart* published 757 articles
on this topic. Valence was found to be slightly positive across all
sources. *Fox News* produced the most total words with the most tokens
on average. This is likely due to the fact *Fox News* transcribes many
of their videos and publishes them in article form. *Politico* featured
the highest number of types on average. *Rush Limbaugh* featured the
lowest readability score on average by grade level while *Slate*
featured the highest grade-level readability score. The large standard
deviations for these statistics, however, preclude conclusions regarding
differences in the sources' use of language, as there is likely a lot of
overlap between sources' use of language.

```{r exp2_descriptive_calculate_kav, include = F, echo = F}
#Kav
final_kav$uniquewords = NA
final_kav$valenceavg = NA
for (i in 1:nrow(final_kav)){

  #add unique
  final_kav$uniquewords[i] = length(unique(unlist(strsplit(final_kav$stemmed[i], " "))))

  #add valence
  tempwords_kav = as.data.frame(unlist(strsplit(final_kav$stemmed[i], " ")))
  colnames(tempwords_kav) = "theword"
  tempwords_kav$tempval = vlookup(tempwords_kav$theword, #index in original data
                    warriner, #where's the stuff you want to put in
                    "V.Mean.Sum", #what column is the stuff you want
                    "Word2") #what's the matching index

  final_kav$valenceavg[i] = mean(tempwords_kav$tempval, na.rm = T)
}
# tapply(final_kav$valenceavg, final_kav$Source, mean)
# tapply(final_kav$valenceavg, final_kav$Source, sd)
# tapply(final_kav$wordcount, final_kav$Source, mean)
# tapply(final_kav$wordcount, final_kav$Source, sd)
# tapply(final_kav$uniquewords, final_kav$Source, mean)
# tapply(final_kav$uniquewords, final_kav$Source, sd)
# tapply(final_kav$Source, final_kav$Source, length)
# tapply(final_kav$wordcount, final_kav$Source, sum)
library(quanteda)
final_kav$FK = NA
for (i in 1:nrow(final_kav)){
  final_kav$FK[i] = unlist(textstat_readability(final_kav$Text[i],
                     measure = "Flesch.Kincaid")$Flesch.Kincaid)
}
# tapply(final_kav$FK, final_kav$Source, mean)
# tapply(final_kav$FK, final_kav$Source, sd)
#Kav Table
####build the table here####
tableprint3 = matrix(NA, nrow = 10, ncol = 11)
colnames(tableprint3) = c("Source", "Mean_Valence", "SD_Valence", "Number_of_Articles", "Total_Words", "Mean_Tokens", "SD_Tokens", "Mean_Types", "SD_Types", "Mean_Readability", "SD_Readability")
#tableprint3[ , 1] = unique(final_kav$Source)
tableprint3[ , 1] = c("The Blaze", "Breitbart", "Fox News", "Sean Hannity",
                     "Huffington Post", "NPR", "New York Times", "Politico",
                     "Rush Limbaugh", "Slate")
tableprint3[ , 2] = tapply(final_kav$valenceavg, final_kav$Source, mean, na.rm = T)
tableprint3[ , 3] = tapply(final_kav$valenceavg, final_kav$Source, sd, na.rm = T)
tableprint3[ , 4] = tapply(final_kav$valenceavg, final_kav$Source, length)
tableprint3[ , 5] = tapply(final_kav$wordcount, final_kav$Source, sum, na.rm = T)
tableprint3[ , 6] = tapply(final_kav$wordcount, final_kav$Source, mean, na.rm = T)
tableprint3[ , 7] = tapply(final_kav$wordcount, final_kav$Source, sd)
tableprint3[ , 8] = tapply(final_kav$uniquewords, final_kav$Source, mean, na.rm = T)
tableprint3[ , 9] = tapply(final_kav$uniquewords, final_kav$Source, sd, na.rm = T)
tableprint3[ , 10] = tapply(final_kav$FK, final_kav$Source, mean, na.rm = T)
tableprint3[ , 11] = tapply(final_kav$FK, final_kav$Source, sd, na.rm = T)
tableprint3[ , c(2,3,6,7,8,9,10,11)] = printnum(as.numeric(tableprint3[ , c(2,3,6,7,8,9,10,11)]), big.mark = "")
tableprint3 = tableprint3[c(5, 6, 7, 8, 10, 1, 2, 3, 4, 9), ]
```

```{r exp2-source-descriptives-kav, results = 'asis', echo = FALSE, message = F}
#change this to match
apa_table(as.data.frame(tableprint3),
          note = "Readability statistics were calculated using the Flesch-Kincaid Grade Level readability formula. V = Valence, T = Tokens or total words, Ty = Types or unique words, FK = Flesch-Kincaid",
          caption = "Kavanaugh - Descriptive Statistics by Source",
          col.names = c("Source", "$M_V$", "$SD_V$", "$N_{Article}$", "$N_{Words}$", "$M_T$", "$SD_T$", "$M_{Ty}$", "$SD_{Ty}$", "$M_{FK}$", "$SD_{FK}$"),
          format = "latex",
          escape = F,
          align = c("l", rep("c", 10)),
          font_size = 'footnotesize',
          placement = "h"
          )
```

Table \@ref(tab:exp2-source-descriptives-gs) displays descriptive
statistics for articles about the partial government shutdown. Like the
Kavanaugh hearing, the sources were similar in average valence (slightly
positive). Once again, there was variation in the number of articles
published by each source on this topic. *Sean Hannity*, *Rush Limbaugh*,
and *The Blaze* published fewer than 100 articles while *Fox News*
published 1,013 articles. *Fox News* again featured the most total words
and mean tokens, but this is likely due to to the presence of a large
amount of video transcriptions that the organization published as
articles. *Politico* had the most types on average. For this topic, *Fox
News* featured the lowest reading grade level while *Slate* featured the
highest reading grade level. For each statistic, the excessively high
standard deviations render any assertions regarding linguistic
differences inconclusive on a descriptive level due to the
aforementioned overlap in sources' language use.

```{r exp2_descriptive_calculate_gs}
#GS
final_gs$uniquewords = NA
final_gs$valenceavg = NA

for (i in 1:nrow(final_gs)){

  #add unique
  final_gs$uniquewords[i] = length(unique(unlist(strsplit(final_gs$stemmed[i], " "))))

  #add valence
  tempwords_gs = as.data.frame(unlist(strsplit(final_gs$stemmed[i], " ")))
  colnames(tempwords_gs) = "theword"
  tempwords_gs$tempval = vlookup(tempwords_gs$theword, #index in original data
                    warriner, #where's the stuff you want to put in
                    "V.Mean.Sum", #what column is the stuff you want
                    "Word2") #what's the matching index

  final_gs$valenceavg[i] = mean(tempwords_gs$tempval, na.rm = T)

}

# tapply(final_gs$valenceavg, final_gs$Source, mean)
# tapply(final_gs$valenceavg, final_gs$Source, sd)
# tapply(final_gs$wordcount, final_gs$Source, mean)
# tapply(final_gs$wordcount, final_gs$Source, sd)
# tapply(final_gs$uniquewords, final_gs$Source, mean)
# tapply(final_gs$uniquewords, final_gs$Source, sd)
# tapply(final_gs$Source, final_gs$Source, length)
# tapply(final_gs$wordcount, final_gs$Source, sum)

library(quanteda)

final_gs$FK = NA

for (i in 1:nrow(final_gs)){
  final_gs$FK[i] = unlist(textstat_readability(final_gs$Text[i],
                     measure = "Flesch.Kincaid")$Flesch.Kincaid)
}

# tapply(final_gs$FK, final_gs$Source, mean)
# tapply(final_gs$FK, final_gs$Source, sd)

#GS Table
####build the table here####
tableprint4 = matrix(NA, nrow = 10, ncol = 11)
colnames(tableprint4) = c("Source", "Mean_Valence", "SD_Valence", "Number_of_Articles", "Total_Words", "Mean_Tokens", "SD_Tokens", "Mean_Types", "SD_Types", "Mean_Readability", "SD_Readability")

#tableprint4[ , 1] = unique(final_gs$Source)
tableprint4[ , 1] = c("The Blaze", "Breitbart", "Fox News", "Sean Hannity",
                     "Huffington Post", "NPR", "New York Times", "Politico",
                     "Rush Limbaugh", "Slate")
tableprint4[ , 2] = tapply(final_gs$valenceavg, final_gs$Source, mean)
tableprint4[ , 3] = tapply(final_gs$valenceavg, final_gs$Source, sd)
tableprint4[ , 4] = tapply(final_gs$valenceavg, final_gs$Source, length)
tableprint4[ , 5] = tapply(final_gs$wordcount, final_gs$Source, sum)
tableprint4[ , 6] = tapply(final_gs$wordcount, final_gs$Source, mean)
tableprint4[ , 7] = tapply(final_gs$wordcount, final_gs$Source, sd)
tableprint4[ , 8] = tapply(final_gs$uniquewords, final_gs$Source, mean)
tableprint4[ , 9] = tapply(final_gs$uniquewords, final_gs$Source, sd)
tableprint4[ , 10] = tapply(final_gs$FK, final_gs$Source, mean)
tableprint4[ , 11] = tapply(final_gs$FK, final_gs$Source, sd)

tableprint4[ , c(2,3,6,7,8,9,10,11)] = printnum(as.numeric(tableprint4[ , c(2,3,6,7,8,9,10,11)]), big.mark = "")

tableprint4 = tableprint4[c(5, 6, 7, 8, 10, 1, 2, 3, 4, 9), ]
```

```{r exp2-source-descriptives-gs, results = 'asis', echo = FALSE, message = F}
#change this to match
apa_table(as.data.frame(tableprint4),
          note = "Readability statistics were calculated using the Flesch-Kincaid Grade Level readability formula. V = Valence, T = Tokens or total words, Ty = Types or unique words, FK = Flesch-Kincaid",
          caption = "Government Shutdown - Descriptive Statistics by Source",
          col.names = c("Source", "$M_V$", "$SD_V$", "$N_{Article}$", "$N_{Words}$", "$M_T$", "$SD_T$", "$M_{Ty}$", "$SD_{Ty}$", "$M_{FK}$", "$SD_{FK}$"),
          format = "latex",
          escape = F,
          align = c("l", rep("c", 10)),
          font_size = 'footnotesize',
          placement = "h"
          )
```